---
title: 'Describing patterns, answering questions'
author: 'Ben Whalley, Paul Sharpe, Sonja Heintz'
# date: "April 2021"
bibliography: [references.bib]
csl: apa.csl
#biblio-style: apa6
link-citations: yes
output:
  webex::html_clean
---

```{r, include=F, echo=F}
source('_first_chunk.R')
```



```{html, child="bs-tab-fix.html"}
```


```{css, child="video.css"}
```



# Overview



```{r, echo=F}
video_data <- list(identifier = "describe-overview", showcode = F, ytidentifier = "yt-embed")
# makermds(video_data)
```

```{r child = '_content/_video_tabs.rmd'}
```



# Central tendency and spread

```{r, echo=F}
video_data <- list(identifier = "describe-centrantend", ytidentifier = "yt-embed")
# makermds(video_data)
```

```{r child = '_content/_video_tabs.rmd'}
```

:::{.exercise}

TODO XXX VERY SIMPLE REVISION EXERCISES IN WORKBOOK
- USING GROUP_BY AND SUMARISE
- MAKING A BOXPLOT WITH A DEMO DATASET

:::


:::{.exercise}

TODO XXX
- in your workbook
- use mutate to create a variable representing weight lost between baseline and end of follow-up in the funimagery data
- make a boxplot of these data

:::


# The *frequency* of responses


```{r, echo=F}
video_data <- list(identifier = "describe-frequency", ytidentifier = "yt-embed")
# makermds(video_data)
```

```{r child = '_content/_video_tabs.rmd'}
```


:::{.exercise}

TODO XXX

- Exercise where they have to do this

```{r}
earnings %>% group_by(education) %>% count()
```


:::


:::{.exercise}

XXX TODO POSSIBLE EXTENSION EXERCISE
THIS COULD BE AN extension activity but low priority for now

```{r}
earnings %>%
  group_by(gender, education) %>%
  count() %>%
  pivot_wider(names_from = "gender", values_from="n")
```

:::





# Relationships



```{r, echo=F}
video_data <- list(identifier = "describe-relationships", ytidentifier = "yt-embed")
# makermds(video_data)
```

```{r child = '_content/_video_tabs.rmd'}
```



:::{.exercisee}

XXX TODO

- What is the correlation between `power` and `engine_size` in the `fuel` dataset?

```{r}
library(corrr)
fuel %>% ggplot(aes(engine_size, power)) + geom_point()
fuel %>% select(power, engine_size) %>% correlate()
```

:::


:::{.exercise}

XXX TODO
- play guess the correlation

:::




# Extension exercises


```{r, child="_content/extension-exercises-premable.rmd"}
```


TODO XXX


:::{.exercise}

XXX TODO possible extension exercise


- Correlations require that there is a *linear relationship* between the two variables
- Which pairs of numeric column in the `fuel` dataset would be suitable for correlation?

```{r, message=F}
# perhaps not mpg and engine_size or power, or engine_size+weight
fuel %>% select(mpg, engine_size, power, weight) %>%
  GGally::ggpairs()
```


:::





# Check your knowledge

Write an answer to each of these questions in the `Check your knowledge` section of your workbook. The answers will be
revealed in Session 5.


TODO XXX



# References
