---
title: 'The immovable mind (Fudōshin)'
author: 'Ben Whalley, Paul Sharpe, Sonja Heintz'
# date: "April 2021"
bibliography: [references.bib]
csl: apa.csl
#biblio-style: apa6
link-citations: yes
output:
  webex::html_clean
---


```{r, include=F, echo=F}
source('_first_chunk.R')
```



<!-- 4.	Evidence 2 – correlations -->
<!-- -	Scatter plot -->
<!-- -	Effect sizes shown in scatter plots (make it a quiz* and/or use http://guessthecorrelation.com ) -->
<!-- -	Bayesian evidence for a correlation -->
<!-- -	Colouring a scatterplot*† -->
<!-- -	Smoothed line plot*† -->




STUFF FROM SESSION 1

We show the benefit of jitter somewhere here now

<!-- he points on the scatter plot are 'jittered' to avoid -->
<!-- overplotting, making the output easier to read. -->

```{r}
mpg %>% 
  ggplot(aes(cty, hwy)) + geom_point()

mpg %>% 
  ggplot(aes(cty, hwy)) + geom_jitter()
```

`r hide('Video script')`
A problem with this plot is that there are many cars with the same number of cylinders. Another way of saying this is
that `cyl` has been plotted using an 'integer (whole number) scale'. Consequently, many of the points are plotted on top
of each other, making the values hard to see. The overplotting issue is easily solved by replacing `geom_point()` with
`geom_jitter()`, which spreads the points out where cars have the same number of cylinders.
`r unhide()`


ALSO SHOW COLOR SCALE PROBLEM (REPEATS FROM PREV SESSION)

```{r}
mtcars %>%
  ggplot(aes(wt, mpg, colour = cyl)) +
  geom_point()
```

The number of cylinders are plotted in continuous shades of blue. This isn't terrible, but it would be clearer if each
number of cylinders had its own colour. This is the same problem we had with some of the other `mtcars` variables. They
have type `dbl`, when they are actually factors. We can use `factor` to fix the colours in the same way we fixed the
x-axis:

```{r}
mtcars %>%
  ggplot(aes(wt, mpg, color = factor(cyl))) +
  geom_point()
```

In this plot it's easier to see that cars with more cylinders tend to be heavier, and have worse fuel consumption.

# Check your knowledge

Write an answer to each of these questions in the `Check your knowledge` section of your workbook. The answers will be
revealed in Session 5.

- What is the difference between `geom_jitter()` and `geom_point()`? 
- Why is `geom_jitter` useful sometimes?


