---
title: 'Quantifying evidence'
author: 'Ben Whalley, Paul Sharpe, Sonja Heintz'
# date: "April 2021"
bibliography: [references.bib]
csl: apa.csl
#biblio-style: apa6
link-citations: yes
output:
  webex::html_clean
---

```{r, include=F, echo=F}
source('_first_chunk.R')
```




```{html, child="bs-tab-fix.html"}
```


```{css, child="video.css"}
```




# Overview

In the [previous workshop](session-4.html) we learned techniques to describe and summarise data. 
In this workshop we take the next step and introduce a technique for quantifying how *sure* we are that the patterns 
we see are real. 

Here we introduce the *Bayes Factor*, or BF. The BF a single number which tells us which of two possibilities are more 
likely, given the data we have collected.

As we'll see, the Bayes Factor can be used as a measure of evidence for both differences in groups and for the relationship between variables (correlation).





# Was it a fluke?

<!-- Summary points -->

- In all data there is *sampling error* or *'noise'*
- We may want to *quantify the evidence* that patterns we see are 'real'
- Bayes Factors quantify evidence by comparing the probability of two scenarios: 
  i) the difference we see is 'real', and
  ii) any difference is just a 'fluke'
- By convention, the larger the Bayes Factor the more likely the pattern is real


<!-- transcript -->

Previously, we emphasised that summaries and descriptions of data are used to *answer research questions*. 
For example, we used a boxplot to compare the differences in the amount of weight participants lost when
treated with either Functional Imagery Training or Motivational Interviewing (the `funimagery` dataset).

```{r, echo=F}
funimagery %>% 
  ggplot(aes(intervention, weight_lost_end_trt)) + 
  geom_boxplot()
```

```{r, echo=F}
fitdiff <- funimagery %>% group_by(intervention) %>% summarise(m=mean(weight_lost_end_trt)) %>% pull(m) %>% diff %>% round(1) %>% abs
```


This plot shows the median and interquartile for each treatment group, and it's immediately obvious 
that --- *in this sample* --- there is a difference between the two groups (`r fitdiff` kg, in fact).

For many purposes this plot is enough. In this case, especially, the difference is quite large and 
we can interpret this plot at face value.

However, researchers often want to *quantify the evidence* in their sample. This is important because in psychology the effects of our interventions are often quite subtle. It *might* be the case that patterns we see in the data are a 'fluke': that is, they have come about by chance. 

It _could_ be the case that if we took another sample from the same population (treated more people with FIT and MI) then we would see no difference.


The Bayes Factor compares two possibilities:

- There really is a difference in how much weight people lost with FIT and MI
- There's NO difference between FIT and MI; any difference we see is likely to be a fluke



-------------



Likewise, we also previously saw how a scatterplot visualises the relation between two columns of data:

```{r, echo=F}
fuel %>% 
  ggplot(aes(engine_size, power)) + geom_point()
```

Again, from this plot the relationship seems quite clear. 

However, we don't have that many data points. The Bayes Factor allows us to quantify how much 
evidence we have for this relationship.

Again, the BF compares two possibilities:

- There is a relationship between `engine_size` and `power`
- There is NO relationship (the pattern we see is just a fluke)


### Signals and noise

We must always remember that, in any dataset we collect, there will be some randomness or chance error. 

Researchers call this sort of randomness "noise". The metaphor here is that this noise it makes it harder for us 
to 'listen' for the true signal in the data.



<!-- No code examples here -->





# Bayesian t test

<!-- SUMMSRY-->


- t-tests allow us to *compare the average score for two groups*
- A Bayes Factor (from a Bayesian t-test) compares two different hypotheses:  H1 (there groups are different) vs. H0 (the groups are identical)
- Use the `ttestBF` function (remember to load the `BayesFactor` package first)
- Large Bayes Factors (e.g. > 10) mean we have a lot of evidence there is a difference between groups
- Very small Bayes Factors (e.g. < .1) mean we have strong evidence there is NO difference
- Bayes factors $>3$ or < $\frac{1}{3}$ provide limited evidence for one of the hypotheses, v.s. the other.
- Bayes factors between $\frac{1}{3}$ and $3$ are inconclusive.




<!-- TRANSCRIPT -->


A 'Bayesian t-test' is a procedure which calculates the Bayes Factor for a difference between groups.

That is, we get a Bayes Factor which tells us how likely the difference we see is to be 'real', rather than a fluke
or just due to sampling variation.


Specifically, the Bayes Fsctor from a Bayesian t-test says how probable it is that two groups have a *different average score* for a *continuous variable*. This probability is *relative* to an alternative hypothesis that the average of each group is actually the same. 

Some people call these the *experimental* and *null* hypotheses, or H1 and H0.


As we saw above, we can look at a boxplot and get an intuition for whether we think the
difference is real by looking at how much the *distributions* of the two groups overlap (e.g. with the boxplot).

[
SHOW BOXPLOT AGAIN USING CODE BELOW
There is not much overlap between the groups, so we intuit the difference is real
]

```{r, eval=F, echo=F}
funimagery %>% 
  ggplot(aes(intervention, weight_lost_end_trt)) + 
  geom_boxplot()
```



If find a large Bayes Factor, this is evidence *for* the experimental hypothesis, as compared with the null hypothesis.


### Running a t-test

To run a t-test and calculate the Bayes Factor we first need to load the `BayesFactor` package:

```{r}
library(psydata)
library(BayesFactor)
```



```{r}
ttestBF(formula = weight_lost_end_trt ~ intervention, data = funimagery)
```

The number $515647.5$ is our Bayes Factor. This is a large number, so we have strong
evidence that the difference in means is real.

We explain how to interpret these numbers in more detail in the section below, but broadly:

- Anything > 3 is evidence *for* a difference
- < 0.33 is evidence *against* a difference
- Anything in between is *inconclusive*




# Bayes Factor for correlation


As we saw above, the Bayes Factor from a 'Bayesian t-test' gives us the evidence two groups are really different.

A Bayes Factor for a correlation is similar, but compares the hypotheses that:

- H1: There is a relationship between the two columns of data
- H2: There is NO relationship (any pattern we see is just due to chance)


Another way to think about it is to imagine we have a set of data points like this:

```{r, echo=F}
set.seed(12)
df <- tibble(y=rnorm(30, 0, 1), x=.75*y+rnorm(30, 0, 1)) %>% mutate_all(scale)
correlate(df)
p1 <- df %>% 
  ggplot(aes(x, y)) +
  # scale_x_continuous(expand = c(0,0) , limits = c(0,2)) +
  # scale_y_continuous(expand = c(0,0) , limits = c(-1,1)) +
  coord_fixed(xlim=c(-2,2), ylim=c(-2,2)) +
  guides(alpha=F) + theme_minimal()

p1 +  geom_point() 
```

We ask: *if we wanted to draw a single straight line through all the datapoints, which would fit the best?* That is, which line would fall the closest to all the data points?

This plot shows some examples of lines we might draw though these data points.
The zero correlation coefficient is drawn in blue. Some examples of other *possible* coefficients are shown in red 
(but remember, it could be anything from -1 to 1).

```{r, echo=F, fig.width=4}
dd <- tibble(s=c(0, -.75,-.25, .25, .75 ), cc=factor(c(1,0,0,0,0)))
dd %>% 
ggplot(aes(x, y)) +
  coord_fixed(xlim=c(-2,2), ylim=c(-2,2)) + 
  geom_point(data=df, aes(x,y), size=.2) +
  guides(color=F) + theme_minimal() +
  geom_abline(data=dd, aes(intercept=0, slope=s, 1 , color=cc)) +
  facet_wrap(~factor(s, levels=unique(s), labels = paste("r = ", s)), ncol=7) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```


The Bayes Factor is estimating how much more likely ANY of the red lines are ($r > 0$ or $r < 0$) than is the blue line (where $r = 0$). 

Only some of these possible red lines are shown in the plot — but the correlation can be anything between -1 and 1. The test is saying, *"given the data we have collected, how much more likely is it that the slope is non-zero, than zero?"*

```{r, echo=F, include=F}
library(BayesFactor)
dfdf <- df %>% as.data.frame()
xyr <- dfdf %>% correlate %>% pull(y)  %>% .[[2]] %>% round(2)
xybf <- correlationBF(dfdf$x, dfdf$y) %>% as_tibble() %>% pull(bf) %>% round()
```

In this case, the correlation coefficient is `r xyr`, and the Bayes Factor for this correlation is `r xybf`.

This means we think it is `r xybf` times more likely that the slope is NOT completely flat, given the data 
we have collected.


### A worked example


Let's say we wanted the Bayes Factor for the correlation between size (carat) and price of 
items in the `diamonds` data set.

The first thing to do is to plot the data, to check that the relationship between these variables
is approximately linear (straight line):

```{r, echo=F}
diamonds %>% 
  ggplot(aes(carat, price)) + 
  geom_point()
```





# Interpreting Bayes Factors

The size of the Bayes Factor is mostly dependent on three things:

- *How large* the effect is (how big is the group difference or how steep is the slope)
- *How clear* the difference is (e.g., how much do the scores from each group overlap)
- *How much data* we have collected (more data lets us be more confident)



:::{.tip}

Bayes Factors can also be influenced by what what we thought was more likely BEFORE we collected
the data. Although, in this course, we always assume that either hypothesis was equally likely, this
doesn't have to be the case. It can be important to incorporate this *prior knowledge*
when calculating Bayes Factors (and it's a big advantage of the approach) but we'll cover 
that in later courses.

:::



### Mechanisms and possible worlds

Another way to think of these alternative worlds (A and B) is as different *hypotheses* about how our data came about.

TODO XXX EXPAND AS ALTERNATIVE EXPLANATION
talk about how we are implying that there is a mechanisms which generates the differences we see...


# Test your knowledge


TODO XXX



# Further reading

- The [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
- The [R Markdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
- Other [R Markdown resources](https://rmarkdown.rstudio.com/)
