---
title: 'Freedom to act (Mushin)'
author: 'Ben Whalley, Paul Sharpe, Sonja Heintz'
# date: "April 2021"
bibliography: [references.bib]
biblio-style: apa6
link-citations: yes
output:
  webex::html_clean
---


```{r, include=F, echo=F}
source('_first_chunk.R')
```




<!-- A literal translation of Mushin is something like 'mind without mind', but in karate and other martial arts the term -->
<!-- denotes the state of mind achieved when free from thoughts of anger, fear, or ego during combat, or in everyday life.  -->

<!-- To achieve mushin takes **repeated practice** and complete familiarity with the technique required to act, such that  -->
<!-- technical details can fade into the background. This leaves the student free to act, responding effortlessly to  -->
<!-- the challenge at hand. -->

We think the analogy to using R is clear:

- If you are anxious, stressed or avoidant you will be distracted
- Getting confident with the basics makes more complex techniques possible 




<!-- - Provide stereotyped examples of the sorts of students positive/negative traits which relate to the theme of session - make it all more concrete -->
<!-- - Provide quotes from previous students illustrating? -->




In this session we cover:

- Loading data from files
- Using simple techniques to answer research questions with data
- Saving intermediate steps using variables


Principles/ideas 

- Using data to answer questions
- Precision and literal-mindedness of R
- Paths and directories


R techniques: 

- Assign variable  (keyword: **"assignment operator"**) and find it in the Env window
<!-- - Explanation of pipes as a way of passing data (an alternative to function arguments) -->
- `read_csv` (from a url)
- the files window; loading a datafile from the same directory as the Rmd file (created by bootstrap script)
- uploading a datafile from your own computer  read_csv it; importance of file extensions
- `filter`, `group_by`, `summarise`, `arrange` (remind them they know how to make scatter and boxplots)
- making a variable containing a subset of a dataframe (sing ` <- ` again)


Activity:

???  set up the teaching as a kind of treasure hunt where students have to find use data to answer questions, gather clues and spell out a word?
Alternatively, just ask them to make LOTs of different plots using the techniques above, not just one:


# Arrange

- "what is the size of the largest diamond (by carat) in the `diamonds` dataset? 
- "what cut were the three largest diamonds in that dataset?"

```{r}
diamonds %>% arrange(-carat) %>% head(3)
```



# Filter

- Introduce operators ` < > = `

```{r}
gapminder::gapminder %>% filter(country=="Kenya")
```

Combined filters

```{r}
gapminder::gapminder %>% 
  filter(country=="Kenya") %>% 
  filter(year > 2000)
```




# Filter and arrange

What was the year Kenyans had the lowest life exp:

```{r}
gapminder::gapminder %>% filter(country=="Kenya") %>% 
  arrange(lifeExp) %>% 
  head(6)
```


What was the highest year? All that changes is the minus sign (reverse sorting)

```{r}
gapminder::gapminder %>% 
  filter(country=="Kenya") %>% 
  arrange(-lifeExp) 

```




# `summarise`

TODO make sure to emphasise giving the summary a name (`average_mpg`)

- This is where we fist encounter the need to give things R-legal names so explan about spaces and special characters
- Emphasise everything should be lower case with underscores

```{r}
mtcars %>% summarise(average_mpg = mean(mpg))
```



# `filter and summarise` together


```{r}
mtcars %>% filter(am==1) %>% summarise(mean(mpg))
```

# `group_by` (with `summarise`)

We might make a plot like this:

```{r}
mtcars %>% 
  ggplot(aes(factor(cyl), mpg)) + 
  geom_boxplot()
```

What if we want these numbers in a table (or to report in our report)?

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(average_mpg = mean(mpg))
```


We can group by two variables at once and get a row for each combination:

```{r}
mtcars %>% group_by(cyl, am) %>% summarise(mean(mpg))
```



# Dividing up graphs

In our previous plot we only showed the difference between cars of 4, 6 and 8 cylinders. We can break the plot down
into multiple panels using '**facets**':


```{r}
mtcars %>% 
  ggplot(aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  facet_wrap(vars(am))   # note to paul... use vars here to avoid quoting of variable names which is confusing/inconsistent with other use
```






# Practice problems

Additional questions

- what country had the highest life expectancy in  1952?


```{r}
gapminder::gapminder %>% filter(year == 1952) %>% 
  arrange(-lifeExp) %>% 
  head(1)
```

- what continent had the highest GDP in 2011?

```{r}
gapminder::gapminder %>% 
  group_by(continent) %>% 
  summarise(average_gdp = mean(gdpPercap)) %>% 
  arrange(-average_gdp)
```



- make a boxplot showing life expectancy by continent

```{r}
gapminder::gapminder %>% 
  filter(year > 2000) %>% 
  ggplot(aes(continent, lifeExp)) + 
  geom_boxplot()

```



# "Mega problem"


Describe these as the 'end of level boss characters'. You need to combine all your skills to beat them...


Make a table which shows the average life expectancy for each continent, sorted from highest to lowest:


```{r}
gapminder::gapminder %>% 
  group_by(continent) %>% 
  summarise(life_expectancy = mean(lifeExp)) %>% 
  arrange(-life_expectancy)
```




# Broken script to fix

- Fix a 'broken' script: Start a NEW R session and make this code work:
 

```{r, eval=F}
liibrary(todyverse)

# make a density plot of of life expectacy with different color lines for each continent
gapminder::gapminder %>% 
  ggplote(aes("lifeExp", colr = "Continent"))  geom_density()

# select only years after 1990
gapminder::gapminder %>% 
  filter(year > 1990)

ggplot(aes(year, lifeExp, color=continent)) + 
  geom_jitter()
  
```


NOTE - we will know all the errors they will see so can provide hints for each of them


Correct version would be: 

```{r}
library(tidyverse)

# make a density plot of of life expectacy with different color lines for each continent
gapminder::gapminder %>% 
  ggplot(aes(lifeExp, color = continent))  + 
  geom_density()

# select only years after 1990
gapminder::gapminder %>% 
  filter(year > 1990) %>%
  ggplot(aes(year, lifeExp, color=continent)) + 
  geom_jitter()
```





