---
title: lifesavR 'Check your knowledge' questions and answers
author: 'Ben Whalley'
output: 
  powerpoint_presentation:
    incremental: true
---

```{r, include=F, echo=F}
source('../website/_first_chunk.R')
```

# Check your knowledge Q and A

## Teachers' notes


Each week we pose 'check your knowledge' questions.

This presentation includes the answers to each question, designed to be used as
a group Q/A exercise in-class.





# Session 1


## How do you run part of a line of R code using the keyboard short cut?

---

Select just that part, and press `ctrl+Enter`



## Which library will you always need to load in your first RMarkdown chunk?

---

`tidyverse`, loaded with `library(tidyverse)` (and often also `library(psydata)`)



## What is `psydata`?

---

A package we provide which includes example datasets.



## How would you look at/inspect a whole dataset?

---

Just type the name of the dataset and 'run' it. In RStudio this provides a data-viewer window.



## What does `glimpse()` do and when is it useful?

---

It provides a preview of all the columns in a dataset, showing the type and first few values of
each.

It's useful when you have a lot of columns in a dataset which would otherwise be
too wide to view on screen.



## What is the 5th column in the `development` dataset?

---

```{r}
psydata::development %>% names %>% .[5] %>% pander::pandoc.verbatim()
```




## Which function makes a plot? (there are many, but we mean the one shown in the worksheet)

---

`ggplot`



## Which function chooses the columns of data used in the plot?

---

The `aes()` function allows us to choose which variables are used for each axis
on the plot.



# Session 2



## What is the `<-` symbol called and what does it do?

---

This is the **assignment operator**. It stores the result from its right-hand side in the variable
named on the left.



## What is the `%>%` symbol called and what does it do?

---

`%>%` is the pipe. 

It passes the data on its left-hand side to the function on the right.



## Which function would you pipe a dataset into if you wanted to see the first few rows?

---

`head`



## Which function is used to select rows from a dataset?

---

`filter`



## How would you select rows where values in a numeric column are between 10 and 20?

---

```
dataset %>% 
  filter(column > 9 & column < 21)
```

If you wrote `dataset %>% filter(column > 10 & column < 20)` this is almost right, but you would be
excluding values of 10 and 20. 

Normally when we say 'between' 10 and 20 we would include 10 and 20.

`filter(column >= 10 & column <= 20)` also works, and works for non-integer values too.



## Which function is used to summarise data?

---

`summarise`



## Which summary function would you use to calculate the mean of a numeric column?

---

`summarise(mean(columnname))`



## How would you calculate the median of a subset of rows in a data frame?

---

    dataset %>%
      filter(SOMETHING HERE) %>%
      summarise(median(columnname))

# Session 3



## Which functions are needed to make a boxplot?

---

As a minimum, in this example we need the `ggplot`, `aes` and `geom_boxplot` functions

<br><br>


```{r}
fuel %>% 
  ggplot(aes(automatic, weight)) + geom_boxplot()
```



## What is the difference between a `dbl` and a `fct` or `ord`?

---

A `dbl` is a 'double precision floating point number'

(that's just 'a number' to you and me).

A `fct` is a 'factor', so a categorical variable. 

An `ord` is an 'ordinal' variable, which is a factor
in which the categories have an ordering.



## Which data types are used for continuous variables?

---

`dbl` or `int` (short for integer).



## Give an example of why the difference between `dbl` and `fct` matters when making a plot (include code examples for this if you can)

---


One example would be this plot: In a boxplot we need the x axis to be a categorical variable, but in
the `fuel` data the `cyl` variable is an integer, so we only get one box rather than 3.

<br><br>

```{r}
fuel %>% 
  ggplot(aes(x=cyl, weight)) + geom_boxplot()
```


---


```{r}
fuel %>% 
  ggplot(aes(x=factor(cyl), weight)) + geom_boxplot()
```



## How can you convert a variable from a `dbl` to a `fct`?

---

```
dataset %>% 
  mutate(myvar_as_factor = factor(myvar))
```



## How could you calculate the mean for one level of a factor?

---

```{r}
development %>% 
  filter(continent=="Africa") %>% 
  summarise(mean(population))
```



## How would you calculate the mean for all levels of a factor (e.g. for continent in the `development` dataset)?

---

```{r}
development %>% 
  group_by(continent) %>% 
  summarise(mean(population))
```



## How would you calculate means and standard deviations for all combinations of levels in two factors?

---

```{r}
development %>% 
  filter(year < 1960) %>% 
  group_by(continent, year) %>% 
  summarise(M=mean(population), 
            SD=sd(population))
```

# Session 4



## What functions in R describe the central tendency of a variable?

---

`mean()` or `median()`.



## What type of column can we use with `mean()`?

---

Numeric (`dbl` or `int`).



## How might you describe patterns in a categorical variable?

---

Making a frequency table (or using a plot).



## What function did we use to create a new column in the `funimagery` dataset?

---

`mutate()`



## In the `funimagery` data, what was the mean change in weight (`kg3 - kg2`) for FIT participants AFTER the end of treatment? (Remember that a negative number means weight was lost.)

---

```{r}
funimagery %>% 
  filter(intervention=="FIT") %>% 
  summarise(mean(kg3 - kg2))
```

The mean change was -2.34 kg: because the number is negative, FIT participants lost (on average) a
further 2.34 kg after the end of treatment.



## What type of plot describes the relationship between two variables?

---

Scatterplot.



## What function do we use to make a plot that shows a relationship between two variables in `ggplot()`?

---

`geom_point()` (or `geom_jitter()`).

# Session 5



## What hypotheses are being compared when we calculate a Bayes Factor for a correlation?

---

H1
:   There is a correlation between x and y

H0
:   There is NO correlation between x and y

The Bayes Factor tells us how many times more likely the data are under H1 than under H0.



## Which function do we recommend to calculate correlation coefficients?

---

`correlate()` in the `corrr` package.



## Which function calculates the Bayes Factor for a correlation?

---

`correlationBF()`



## Which function calculates the Bayes Factor for a difference in the average of two groups?

---

`ttestBF()`



## What direction slope does a correlation of -0.5 imply? What would the scatterplot look like?

---

A negative relationship: the slope goes down. As x increases, y tends to decrease, so the points
slope downwards from left to right.



## What conclusion should we draw if the Bayes Factor for a correlation is 2.1? How about if it was 0.21, or 210?

---

-   2.1 is inconclusive (it is between 1/3 and 3)

-   0.21 is moderate evidence against the correlation (i.e. for the null hypothesis), because
    $1 / 0.21 = 4.76$: the data are about 4.8 times more likely under H0 than H1

-   210 is strong evidence in favour of the correlation



## What conclusion should we draw if the Bayes Factor for a t test is 0.8? How about if it were > 10?

---

-   0.8 is inconclusive

-   \>10 is strong evidence for a difference



## What is the difference between `ttestBF()` and `correlationBF()`?

---

`ttestBF()` calculates the BF for a difference between two groups; `correlationBF()` calculates the
BF for a correlation between two continuous variables.

# Session 6



## What is an intercept?

---

The expected value of the outcome when the predictor is equal to zero.



## What is the slope?

---

The increase in the outcome for a unit change in the predictor.



## What is a residual?

---

The distance between an observed outcome value and the value predicted by the best fit line.



## Which function fits a regression model?

---

`lm()`



## Which function calculates residuals and predictions from a saved `lm` model?

---

`augment()` (from the `broom` package). It adds the `.fitted` and `.resid` columns.



## What are the three assumptions of regression introduced in this session?

---

1. The 'true' best-fit line is straight (i.e. _not_ curved)
2. Model residuals are approximately normal
3. The residuals are of equal size along the regression line

# Session 7



## We prefer models with smaller or larger residuals?

---

Smaller.



## The $R^2$ statistic ranges from ...?

---

0 to 1.



## It's better for a model to have a smaller or larger $R^2$ value?

---

Larger.



## If we calculate the Bayes Factor for `modelA / modelB` and the result is > 1, this provides evidence that ...?

---

We have more evidence for model A than model B: the data are more likely under model A than under
model B.



## A Bayes Factor of 25 means we have ... evidence?

---

Strong evidence.



## A Bayes Factor of 0.05 means we have ...?

---

Strong evidence for the other hypothesis (e.g. the null).

Taking the reciprocal, $1 / 0.05 = 20$: the data are 20 times more likely under the other
hypothesis.
