With categorical data it's not possible to calculate averages; what would the
average of 3 apples and 4 oranges be?

Nonetheless, it is still useful to be able to describe the data and spot patterns.

For example, if we were thinking about the `earnings` dataset we might
ask things like:

- *"How many men and women were included in the sample?"*
- *"What proportion of participants worked in the public sector?"*
- *"Was this proportion similar for men and women?*"


The earnings data contained variables called `gender` and `job`, which give us the
information to answer these questions:

```{r}
earnings %>% glimpse
```



### Using `count()`


To count the frequency of different responses we can use the `count` function.



For example, the following statement counts the total number of responses in our sample:


```{r}
# this gives us the total number of rows
earnings %>% count()
```

However, we often want to count the number of observations by-group. For example, if we write:

```{r}
# numbers of men and women
earnings %>%
  count(gender)
```

Here we added `gender` inside `count()` so R knows to count the responses from each group separately.


We can do arithmetic on these numbers to get the percentage of women in the sample:

```{r}
2353 / 4483 * 100
```

Later on you'll find ways to create a new column containing percentages, but this is fine for now.


### Multiple subgroups

What's neat about `count()` is that we can add more than one variable at once, and we get a count for each subgrouping:

```{r}
earnings %>%
  count(job, gender)
```

You can re-order the table by changing the order in which you include `gender` and `job`.


