
With categorical data it's not possible to calculate things like an average;
what would the average of 3 apples and 4 oranges be, for example?

Nonetheless, it is still useful to be able to describe the data and spot patterns.

For example, if we were thinking about the `earnings` dataset we might
ask things like:

- *"How many men and women were included in the sample?"*
- *"What proportion of participants worked in the public sector?"*
- *"Was this proportion similar for men and women?*"


The earnings data contained variables called `gender` and `job`, which give us
information to answer these questions:


```{r}
earnings %>% glimpse
```


### Using `count()`


To count the frequency of different responses we can use the `count` function, which
is again from the tidyverse package. [load tidyverse now]

For example, the following statement counts the total number of responses in our sample:


```{r}
# this gives us the total number of rows
earnings %>% count()
```

However, we often want to count the number of observations by-group. For example, if we write:

```{r}
# numbers of men and women
earnings %>%
  count(gender)
```

Here we added `gender` inside `count()` so R knows to count the responses from each group separately.


Note, this does the same thing as writing this, although one line shorter:

```{r}
earnings %>%
  group_by(gender) %>%
  count()
```



We can do arithmetic on these numbers to get the percentage of women in the sample:

```{r}
2353 / 4483 * 100
```

Later on you'll find ways to create a new column containing percentages (using mutate), but this is fine for now.


### Multiple subgroups

What's neat about `count()` is that we can add more than one variable at once, and we get a count for each sub-grouping that is formed. For example:

```{r}
earnings %>%
  count(job, gender)
```

You can re-order the table by changing the order in which you include `gender` and `job`.



### Making a plot

If you want to make a plot of counts, then `stat_count` can be useful:

```{r}
earnings %>%
  ggplot(aes(gender)) +
  stat_count()
```
