In this video we'll use the `funimagery` dataset (in the `psydata`) package.
This contains data from a randomised controlled trial run in
Plymouth [@solbrig2019functional].

The study compared two treatments for weight loss: Functional
Imagery Training (FIT) and Motivational Interviewing (MI).


We can see the columns in the dataset using `glimpse()`:

```{r}
funimagery %>% glimpse
```

Weight was measured three times, and these measurements are in the columns called `kg1`, `kg2` and `kg3`.

- The first (`kg1`) was taken at *baseline*, when participants joined the study
- The second (`kg2`) was taken at the end of treatment, 6 months after baseline
- The final measurement (`kg3`) was made 12 months after baseline

The `weight_lost_end_trt` column is the *difference* between `kg1` and `kg2` -- i.e.
how much weight the participant lost after their treatment.



We have already seen a plot like this which can compare scores between categories:

```{r}
funimagery %>%
  ggplot(aes(intervention, weight_lost_end_trt)) +
  geom_boxplot()
```


This graph is helpful, and makes the differences really clear. But what if we want the actual numbers in a table, or to report in the text of our paper?



### Using `filter()` and `summarise()`

One option would be to filter our data first and then summarise each subset in turn:

```{r}
funimagery %>%
  filter(intervention=="MI") %>%
  summarise(mean(weight_lost_end_trt))
```

The output (underneath the code) shows us the mean for the MI group. We then have to repeat that step for the FIT group:

```{r}
funimagery %>%
  filter(intervention=="FIT") %>%
  summarise(mean(weight_lost_end_trt))
```


That would be repetitive though, and would get annoying if there were lots of categories. Imagine doing something
similar for each continent in the `development` data, for example.


### Using `group_by()`

Instead of using `filter()`, we can use `group_by()` to split our dataset into multiple groups, summarising each one separately:


```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean(weight_lost_end_trt))
```


Using `group_by()` before `summarise()` splits up the data depending on which `intervention` the participant
was randomised to.  The `summarise()` function calculates the mean for each group separately.

The result is a new `data.frame` with two columns: The name of the `intervention`, and the average weight lost. We could save
this in a new variable for use later, if we liked:

```{r}
average_weight_losses <- funimagery %>%
                          group_by(intervention) %>%
                          summarise(mean(weight_lost_end_trt))
```

If you run this code, you can now see the new variable in the **Environment**, so it's stored for later.



### Nested groups

Sometimes we might want to group by more than one variable at once.

Imagine we wanted to see if men and women responded differently to the two treatments?

With `group_by()` this is really easy: just add the name of another column to group on (`gender`), separated with a comma:

```{r}
funimagery %>%
  group_by(gender, intervention) %>%
  summarise(mean(weight_lost_end_trt))
```


It looks like women lost a bit more weight on average, but FIT was better than MI for both men and women.



### Multiple statistics

With `summarise()` we can calculate multiple statistics at once for each of the groups:

```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean(weight_lost_end_trt), sd(weight_lost_end_trt))
```

### Give your new variables a name

R gives the new columns a name based on the function we use to `summarise()` them.

For example, if we use `mean()` on the `weight_lost_end_trt` variable then the new column is called `mean(weight_lost_end_trt)`.


:::{.direction}

- rerun code above and highlight where the name of the variable comes from

:::

Unfortunately, these new variable names contain parentheses and (sometimes) spaces which makes them harder to work
with. It's much better to give the new column a specific name like this:

```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean_weight_lost_end_trt = mean(weight_lost_end_trt), sd_weight_lost_end_trt = sd(weight_lost_end_trt))
```

As a reminder, the names you use shouldn't include spaces or other 'special' characters.
