
As we saw before, scatterplots show the relationship between two variables.

Using the `mpg` data, we could plot the size of car engines (`displ`, short for displacement) against their fuel efficiency on
the highway in miles per gallon (`hwy`):

```{r, eval=F}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point()
```

:::{.direction}
* Run code to generate plot
:::

*NOTE: This first plot is not shown here to save space, but is shown in the video.*

We can add colour to distinguish the points in this plot — for example, what kind of car each point represents.

So far, you've used the `aes()` function to define which variable is plotted on the x and y axes.  The `aes()` function is short for 'aesthetics' (what we can see) and *connects columns in our data to visual aspects of a plot*, like the x and y axes.

The `colour=...` option adds to this, and says which column is used to colour the points.

In the `mpg` data, the `drv` column contains one of three values. The value categorises cars based on which wheels are
driven the engine. An `f` in `drv` means a front wheel drive car, `r` means rear wheel drive, and `4` means four
wheel drive.

We can write `colour=drv` to tell R to colour each point, depending on which wheels are driven:

:::{.direction}
1. Add `colour=drv`
1. Run code to generate plot
:::

```{r, eval=T}
# colour each point; different colour for each type of transmission
# front, rear and four wheel drive
mpg %>%
  ggplot(aes(displ, hwy, colour=drv)) +
  geom_point()
```

Notice that `ggplot()` automatically adds a key to the plot, mapping the colour of the points to the categories in `drv`.

This is an example of using a *categorical* variable to colour our points.
