Patterns within a dataset are often easier to see after the data is summarised.

Summarising data means reducing lots of numbers to a smaller number.

For example, the mean is a summary which tells us about the central tendency.

To summarise data in R, we use `summarise()`, with a "summary function" which reduces some data values.

For example, the `mean()` summary function calculates the mean of some data values, in this case the `mpg` column:

```{r}
fuel %>%
  summarise(mean_mpg = mean(mpg))
```
The result is a `data.frame` with a single column named `mean_mpg` (`mean_mpg =` defined the column name).

If you don't assign the result of the summary function to a column name, it is named after the summary function:

```{r}
fuel %>%
  summarise(mean(mpg))
```

You should normally specify a column name, because columns named after summary functions
are awkward to process later in a pipeline.

The mean is just one of many summary functions.

We could also summarise `mpg` using a different measure of central tendency, the median.

In this case we would use the `median()` summary function:

```{r}
fuel %>%
  summarise(median_mpg = median(mpg))
```

Another common summary is the standard deviation, which tells us how far, on average, data values are from the mean.

We can calculate the standard deviation using the `sd()` summary function:

```{r}
fuel %>%
  summarise(sd_mpg = sd(mpg))
```

R only allows certain characters for column (and variable) names.

Errors caused by invalid characters can be confusing, but they're easy to avoid if you follow these simple rules:

1. Only use lower case letters from the alphabet (a-z)
1. Use underscores `_` instead of spaces

You can also use `summarise()` with more than one summary function.

Seperate each summary function with a comma:

:::{.direction}


- Highlight comma in `summarise()`.

:::


```{r}
fuel_summary <- fuel %>%
  summarise(m = mean(mpg), sd = sd(mpg))
```

:::{.direction}

- Show `fuel_summary` in **Environment**

:::

Remember, the out from `summarise()` (in this case `fuel_summary`) is a `data.frame` that you could do more processing
on in extra steps if needed.

### Using `filter` with `summarise`

One powerful thing about R and the `tidyverse` is that you can combine the techniques
you learn to make more sophisticated analyses.

A simple example of this is using `filter()` with `summarise()`.

For example, if we want to calculate the average life expectancy of only Asian countries
in the `development` dataset we can write:

```{r}
development %>%
  filter(continent=="Asia") %>%
  summarise(asian_life_expectancy = mean(life_expectancy))
```
