
Each instruction we give R produces a result.

By default, these results are shown below the code chunk.

Sometimes, though, we'd like to save the results. Perhaps we want to use them in another
calculation later?

To do this we can create a **variable**.

A variable is just a *container*.  We can store things in it: the results of simple
calculations, datasets, tables, plots, or even statistical models.

To make variables we use the assignment operator, which looks like this `<-`.

It looks like an arrow that points to the left, but it's made up of an angle bracket and a hyphen,
placed next to each other.

The arrow is a reminder that the value on the right hand
side will be assigned (stored) in the variable named on the left hand side.


For example, we can write this:

```{r}
# assign the current year to a variable
current_year <- 2021
```

And we now have a variable called `current_year` which contains the number `2021`.


The simplest way to check what a variable contains is to write it's name, and then
run that line of code:

```{r}
# print the variable we created to the console window
current_year
```


You might wonder *where* these variables get saved. In most cases, variables you create are stored in what's called the
`Global Environment`. You can see them in the **Environment** pane in RStudio.

The names of variables are shown in the column on the right, and their values in the column on the left.

[HIGHLIGHT NAME AND VALUE OF A VARIABLE IN THE ENVIRONMENT.]

---------------


Once stored, we can use variables as the *input* to new calculations.

For example, in this example we are adding 1 to the current year and storing the result in another
variable called `next_year`:


```{r}
# do a calculation to create a new variable from an existing one
next_year <- current_year + 1
```


--------------------

Naming variables sensibly is more important than you might realise.

You will save yourself a lot of trouble if you give your variables short names which describe the data they contain. Use the underscore `_` if you need to use more than one word e.g. `current_year`.

Variable names can't contain spaces, and it's best to avoid special characters or capital letters.
