---
title: "LifesaveR: Session 2"
output: html_document
---


```{r}
# use this chunk to load all the packages you need
library(...)
```


# lifesaveR: Workbook 2

This R Markdown document is your workbook for session two of the lifesaveR module.


# Storing data in variables

## Exercise 1

Run this chunk:

```{r, eval=F}
# assign the current year to a variable
current_year <- 2021

# print the variable we created
# (it should display below the chunk, or in the console pane)
current_year

# do a calculation to create a new variable from an existing one
next_year <- current_year + 1
```

## Exercise 2

1. Create a new chunk.
2. Assign the results of the calculation `2 * 35` to the variable `seventy`.
3. Run the chunk.

## Exercise 3

1. Create a new chunk.
2. Use R to calculate your age in the year 2051.
3. Save the result in a variable _with a descriptive name_.

# Passing data with pipes `%>%`

## Exercise 4

1. Create a new chunk.
2. Load the `tidyverse` and `psydata` libraries.
3. Pipe the `fuel` dataset into `head()` and assign the results to a variable called `fuel_head`.
4. Use the **Environment** pane to open `fuel_head`.

# Loading data from elsewhere

## Exercise 5

1. Use your web browser to download
[https://raw.githubusercontent.com/benwhalley/lifesavR/main/website/data/world-happiness-report-2021.csv](https://raw.githubusercontent.com/benwhalley/lifesavR/main/website/data/world-happiness-report-2021.csv) to your computer.
2. Upload `world-happiness-report-2021.csv` to the server.
3. Create a new chunk.
4. Read `world-happiness-report-2021.csv` into a variable with a descriptive name.

# Selecting rows with `filter()`

## Exercise 6

Filter the `development` dataset to show countries with a population greater than 100 million.

## Exercise 7

Show countries with a population greater than 100 million and life expectancy greater than 70.


# Summarising data using `summarise()`

## Exercise 8

1. Copy the code in the worksheet which calculates the median miles per gallon.
2. Paste the code into a new chunk.
3. Amend the code to calculate the median weight.
4. Note that `fuel` stores `weight` in kilos.

## Exercise 9

Amend the code in Exercise 8 to also calculate the mean and standard deviation of `weight`.

## Exercise 10

Using the `fuel` dataset, `filter()` and `summarise()`, calculate:

- the mean miles per gallon of cars with engine size > 3000
- the standard deviation of miles per gallon in cars with 6 cylinders



# How to fix code errors

## Exercise 11

1. Select the R Studio menu option `Session > Restart R`.
2. Ensure that the **Environment** pane is empty (see worksheet for how to do this).
3. Correct all of the errors in the chunk below.

```{r}
libary(Tidyverse)
library(psydata

# select cars with less than 8 cylinders
development %>%
  filter(cyl > 8)

# calculate the average life expectancy in Africa
development
  filter(continent = "africa") %>%
  summarise(average(life_expectancy))

# scatterplot of life expectancy against year with continent in colour
ggplot(aes(year, lifeExp, colour=continent)) %>%
  geom_scatterplot()
```



# Check your knowledge

Write your answers below each question and save your workbook for later.

## What is the `<-` symbol called and what does it do?

## What is the `%>%` symbol called and what does it do?

## Which function would you pipe a dataset into if you wanted to see the first few rows?

## Which function is used to load data from a CSV file?

## What must you do before you can load data on the RStudio server?

## Which function is used to select rows from a dataset?

## How would you select rows where values in a numeric column are between 10 and 20?

## Which function is used to summarise data?

## Which summary function would you use to calculate the mean of a numeric column?

## How would you calculate the median of a subset of rows in a `data.frame`?

# Extension exercises

## Extension exercise 1

In a new chunk, write the R code to produce the `data.frame` shown in the worksheet.

## Extension exercise 2

Filter and summarise the `iris` dataset in order to answer the question in the worksheet.