A 'Bayesian t-test' is a procedure which calculates the Bayes Factor for a difference between groups.

That is, we get a Bayes Factor which tells us how likely the difference we see is to be 'real', rather than a fluke or due to sampling variation.

Specifically, the Bayes Factor from a Bayesian t-test says how probable it is that two groups have a *different average score* for a *continuous variable*. This probability is *relative* to an alternative hypothesis that the average of each group is actually the same.

Some people call these the *experimental* and *null* hypotheses, or H1 and H0.


As we saw above, we can look at a boxplot and get an intuition for whether we think the
difference is real by looking at how much the *distributions* of the two groups overlap (e.g. with the boxplot).

:::{.direction}

Show boxplot again using code below

:::

There is not much overlap between the groups, so we intuit the difference is real



```{r, eval=F, echo=F}
funimagery %>%
  ggplot(aes(intervention, weight_lost_end_trt)) +
  geom_boxplot()
```



If find a large Bayes Factor, this is evidence *for* the experimental hypothesis, as compared with the null hypothesis.


### Running a t-test

To run a t-test and calculate the Bayes Factor we first need to load the `BayesFactor` package:

```{r}
library(psydata)
library(BayesFactor)
```


And then use the `ttestBF` function.


```{r}
with(funimagery,  ttestBF(weight_lost_end_trt, intervention))
```

```{r}
fitbf <- with(funimagery,  ttestBF(weight_lost_end_trt, intervention)) %>%
  as_tibble() %>% pull(bf) %>% .[1]
```


The number `r fitbf %>% sprintf("%.6g", .)` is our Bayes Factor. This is a very large number written in scientific notation.

If we wrote it out in full, it would be `r fitbf %>% sprintf("%.0f", .)`.

So, we have strong evidence that the difference in means is real.

We explain how to interpret Bayes Factors in more detail in the section below, but large numbers are evidence *for* a difference.
