Boxplots are useful for visualising differences between categories or groups.

Using `ggplot`, making a boxplot is similar to a scatterplot.

If we haven't already done it, we'd need to load the `tidyverse`:

```{r, echo=T, warning=F, message=F}
library(tidyverse)
```


------------------------

To make the plot, we first choose the data we want to plot and use a pipe.

```{r, eval = F}
iris %>% 
```


Next we use `aes(x = ..., y = ...)` to select the columns we want to use in the data.
This time we have chosen `Species` for the x axis, because it is a *categorical* variable

```{r}
iris %>% 
  ggplot(aes(x = Species, y = Petal.Length)) 
```



[demonstrate adding x and y variables using auto complete in RStudio]

You'll notice that I used the RStudio autocomplete feature to write variable names.
This makes things much easier — if you type variables by hand it's easy to make spelling errors
or typos and this leads to errors in R.


If we run that code, `ggplot` draws the plot axes, but doesn't add the data yet:

[run the code so far]


To finish, we add `geom_boxplot()` and actually draw the boxes:

```{r}
iris %>% 
  ggplot(aes(x = Species, y = Petal.Length)) + 
  geom_boxplot()
```


Now we have a boxplot, with one box drawn for each value of `Species`.



### Interpreting boxplots

In a boxplot: 

- the thick line is the median. 

- the height of the box indicates the inter-quartile range (IQR). 

- the whiskers vary a bit, depending on the software you use — but normally show the range/spread of the data


The IQR contains the middle 50% of the data. A wider IQR indicates greater variation (spread) in a dataset.

The vertical lines connected to the top and bottom of the box are called 'whiskers', and
give some indication of the range of the data. 

Exactly what the whiskers show differs depending on the particular options you use to draw a boxplot. 

The default is to show the points that are no more (or less) than 1.5 times the IQR above (or below) the top (bottom) of the box. 

In this version of a boxplot, any data point outside the range of the whiskers is described as an 'outlying point' and is shown individually as a dot.
