
Datasets contain are like spreadsheets: they are organised into columns and rows.

In R, datasets are normally stored in a container called a `data.frame`. They can also be stored in a `tibble` (these are basically the same thing).



### Columns

Each column in a dataset has a name.

We sometimes call the columns *variables*, because each column will often relate to
a variable in our study.

However, this can be a bit confusing because --- in R --- variables can actually contain whole datasets.
`fuel`, for example, is the name of a variable which contains an example dataset, provided by the `psydata` package.

[SHOW LIBRARY(PSYDATA) AND THEN THE FUEL DATASET]

But these words are used flexibly and interchangeably, so we'll just have to get used to it.
It's normally clear which type of variable we mean from the context.


### Rows

Each row in a dataset represents an *observation*.

In different datasets an observation might correspond to an individual participant, a whole country, or even just a single button press in an experiment.



### Packaged datasets

Some datasets are built into R packages as examples for beginners.

For this course, we created a package called [`psydata`](https://github.com/benwhalley/psydata/)
which includes the data we need for teaching.

This is installed on the RStudio server. To load it we run:

```{r}
library(psydata)
```

We can see from the loading message that one of the datasets is called `fuel`. This contains data about cars --- things like weight, fuel economy, engine size.

Let's display this data in using a new chunk.  If we type the word `fuel`, select this variable name with our cursor, and 'execute' it, we can see the data it contains:

```{r}
fuel
```


By default this shows the first ten rows and columns of the data. You can see other rows using the `Next`,
`Previous` and number buttons below the data.

If your browser window is very narrow you may need to view some of the columns by using the arrow next to the final, right-hand column.



You can get information about the columns in all these example datasets by typing:
`help(name_of_the_dataset_you_want_to_know_about)`. For example:

```{r}
help(fuel)
```




### Exploring and checking data

There are two ways we recommend to inspect and check data you are using.

1. Typing the name of the dataset, and running that as code
2. The `glimpse()` function, which shows a list of all the columns and some of the data
3. The `head()` function, which shows the first 10 rows



To use glimpse:

```{r}
glimpse(fuel)
```

This shows a list of all the columns in the dataset, the type of data stored in each column, and as many observations (datapoonts) as will fit on a single line.

`glimpse` is a really useful view to check which columns are available in a dataset before using them.



Using head:

```{r}
head(fuel)
```

This prints the first 10 rows of all the columns.   `head()` is really useful for checking the actual datapoints are as you expect before using them.





### Why are we talking about cars not psychology?

In this course we mostly use very simple datasets, and some of them aren't even about psychology.

Some students ask why we don't always use psychological examples. If this hasn't troubled
you then you could skip to the next section, but we thought we should explain:


We think the `fuel` dataset (and others, like `iris`, and `development`) have a number of benefits.

First, they are either built into R, loaded in common packages, or available in the `psydata` package.
This makes them easily available for everyone.

Second, these data relate to concrete, easy to understand phenemena (e.g. weight, length, number of gears).
This means you don't have to hold in mind any complex psychological/theoretical ideas for the examples to make sense.

Third, the relationships in these datasets are clear, and there aren't too many data points. Real data are
often more messy because many psychological constructs are hard to measure.

Our experience is that, when learning R, it pays to keep everything as simple as it possibly can be.
The skills and concepts involved in analysing these data are the same though.

R --- and the techniques and statistics we teach --- are used right across the natural sciences

[ TOOD show examples of plots and analyses here]

If you're still not convinced — don't worry ... we do include some clinical examples, and we will
be collecting our own psychological data soon enough and analysing that.
