
Datasets contain are like spreadsheets: they are organised into columns and rows.


### Columns

Each column in a dataset has a name.

We sometimes call the columns *variables*, because each column will often relate to
a variable in our study.

However, this can be a bit confusing because --- in R --- variables can actually contain whole datasets.
`mtcars`, for example, is the name of a variable which contains an example dataset.

[show mtcars]

But these words are used flexibly and interchangeably, so we'll just have to get used to it.
It's normally clear which type of variable we mean from the context.


### Rows

Each row in a dataset represents an *observation*. In different datasets an observation might correspond to an individual participant, a whole country, or even just a single response in an experiment made by a single person.


### Differet names

In R datasets (columns/rows of data) are stored in a special type of variable called a
`data.frame`.  

You will also see references to datasets as `tibbles`.

A `tibble` is just a special type of `data.frame`, so you can treat these the same.

All these terms are mostly interchangeable:
- dataset
- Dataframe or `data.frame`
- `tibble`


### Built in datasets

There are quite a few data sets which are built into R to provide examples for beginners.

One of these built in datasets is called `mtcars`. This contains data about cars that was published in a US
magazine called Motor Trend.

Let's display this data in using a new chunk. As we did with the `diamonds` tibble, if we
type `mtcars`, select the variable name and execute it, we can see the data it contains.

```{r}
mtcars
```


By default this shows the first ten rows and columns of the data. You can see other rows using the `Next`,
`Previous` and number buttons below the data. You can see additional columns using the arrow next to the final,
right-hand column.


Other built in datasets:

- the diamonds dataset is loaded with tidyverse (when you run `library(tidyverse)`)

- The `gapminder` package loads a dataset which is also called `gapminder`. This contains data about life expectancy, GDP per capita, and population by country.

We can load the gapminder data the same way we loaded the `diamonds` dataset.
We load the `gapminder` package first, type the name of the dataset (also
`gapminder`) and run it.

Again, we can use the navigation buttons to explore the data.


```{r}
library(gapminder)

gapminder
```





### Exploring and checking data


The `head()` function allows you to explore the first few rows of a dataset. The code `mtcars %>% head()` prints the
first six rows of `mtcars`. The `%>%` function is called the 'pipe'. We'll explain more about how it works in the next
session. For now, all you need to know is that the function on the right hand side of the pipe (`head()`) is applied to
whatever is on the left hand side (in this case, the `mtcars` dataset).

The `glimpse()` function is useful for exploring datasets with lots of columns, as it allows you to see all columns at
once. Using `glimpse()` we can have the columns of a dataset run down the page, and data run across.

The [World Happiness Report](https://www.kaggle.com/ajaypalsinghlo/world-happiness-report-2021) is a survey of the state
of global happiness. Running the code `happiness %>% glimpse()` shows all of the columns. This is like rotating the
output you saw earlier anti-clockwise by 90 degrees. The code displays all of the columns as rows, and as many
observations from the dataset as will fit on a single line.

Each column in a `data.frame` has an associated type. The second column of the `glimpse` output shows you the type of each column. This dataset includes columns with two types:

* `dbl` is short for 'double-precision number', a number with one or more decimal places
* `chr` — short for 'character', a variable which contains text (e.g. an email address)

Other types include:

* `fct` is short for 'factor', a categorical variable (e.g. a specific response to a multiple-choice question)
* `int` is short for 'integer', a variable which contains whole numbers (e.g. a participant id number), and
* `ord` — short for 'ordered'; a variant of *fct* where the categories have a particular order (e.g. responses like 'Wost' < 'Better' < 'Best')

The importance of knowing the type associated with a column will become clear in a later session.
