If we have a single numeric variable we can describe:

1. the *central tendency* of the data: e.g. mean, median
2. the *spread* or distribution of the data: e.g. the standard deviation or interquartile range

Even simple descriptive statistics like the mean or standard deviation enable us to *answer research questions*. For example, if we consider the `funimagery` data describing the RCT of functional imagery training, we could ask:

- *"what was the typical weight of participants at baseline?"* or
- *"what was the **range** in which most participants' weight fell?"*

You have already seen how techniques like `group_by()` and `summarise()`, or graphs like boxplots, can help
calculate and present these *descriptive statistics*.


```{r}
# (this is a recap of earlier material)
# typical weight at baseline
funimagery %>%
  summarise(mean(kg1))


# a boxplot showing the IQR as the box. The IQR includes 50% of participants
# so, we can see 50% of participants weighed between 80 and 100kg at baseline
funimagery %>%
  ggplot(aes(intervention, kg1)) +
  geom_boxplot() +
  scale_y_continuous(n.breaks = 10) # this extra line just adds more marks on the y axis
```




## Describing *differences*

The previous table and boxplot showed patients' weights at the start of the study.

There is also a variable in this dataset called `weight_lost_end_trt`, which shows how much weight people lost
between starting and completing FIT or MI. In a previous session we made a boxplot like this:

```{r, echo=F}
funimagery %>%
  ggplot(aes(intervention, weight_lost_end_trt)) +
  geom_boxplot()
```

However, in clinical trials, it's important to measure participants for longer periods to judge whether the effect of a treatment is sustained. Interventions for obesity and overweight can be successful, but patients may later regain weight [@hall2018maintenance]. Estimating how long weight loss is sustained is important because it changes the long term prognosis of patients, and how cost-effective an intervention is.

The `funimagery` data come from a study which followed people for 6 months after completing treatment
(12 months after joining the study). The `kg1` column records weights at baseline, and the`kg3` columns
records observations made at the end of follow-up


This means we can calculate weight loss from baseline to follow-up (not just the end of treatment, which has already been done for us).

We need to create a *new column* in our dataset. Let's call this `weight_lost_end_followup`.

To calculate this new column we need to subtract weight at baseline (`kg1`) from weight at the end of follow-up (`kg3`).

In R, we can do this with the `mutate` function:

```{r, eval=T}
# use `mutate` to create a NEW COLUMN of data
# this code shows the result just below the code chunk
funimagery %>%
  mutate(weight_lost_end_followup = kg3 - kg1)
```

[RUN THE CODE ABOVE AND SHOW STUDENTS THE RESULT. POINT OUT THAT THIS HAS NOT BEEN STORED
ANYWHERE — JUST DISPLAYED IN THE RSTUDIO GUI BELOW THE CODE CHUNK]


We almost certainly want to STORE this extra column so we can use it again.
To do this we create a *copy* of the whole `funimagery` dataset by *assigning*
the result of `mutate())`

The assignment operator is the left hand arrow, `<-`:

```{r}
funimagery.edited <- funimagery %>%
  mutate(weight_lost_end_followup = kg3 - kg1)
```


[SHOW STUDENTS NEW VARIABLE HAS BEEN CREATED IN THE ENVIRONMENT WINDOW]

So the code above:

- takes the `funimagery` data and pipes it to the `mutate()` function, which
- adds a new column, called `weight_lost_end_followup`
- this new column is made by subtracting `kg1` (baseline) from `kg3` (end of followup)
- stores the new dataset (with the extra column) in a new *variable* called `funimagery.edited`:



We can then use this new variable to do more work, like making a new boxplot:

```{r}
# boxploot of weight lost at end of follow-up using new column
funimagery.edited %>%
  ggplot(aes(intervention, weight_lost_end_followup)) +
  geom_boxplot()
```

If anything, it looks like the difference between groups is even BIGGER after follow-up than it was at the end of treatment, which is very promising.
