With categorical data it's not possible to calculate averages, but it is still useful to be able to describe the data. For example, if we were thinking about the `earnings` dataset we might ask:

- *"How many men and women were recruited?"*
- *"What proportion of participants worked in the public sector?"* 
- *"Was this proportion similar for men and women?*"


The earnings data contained variables called `gender` and `job` which can answer these questions:

```{r}
earnings %>% glimpse
```


To count the frequency of different responses we can use the `count` function.


For example: XXX TODO EXPLAIN THIS CODE AND WRITE AN EXERCISE


```{r}
# this gives us the total number of rows
earnings %>% count()

# numbers of men an women
earnings %>%
  group_by(gender) %>%
  count()

# later we'll show ways of doing arithmetic on the results, but for now
# if we simply divide men/total we get the proportion/percentage of men
2353 / 4483 * 100



# we can do the same for job:
earnings %>%
  group_by(job) %>%
  count()

# proportion in the private sector
3447 / 4483 * 100


# we can use group_by again to get the split for men and women:
earnings %>%
  group_by(job, gender) %>%
  count()

1940
```

