```{r, eval=F}
liibrary(todyverse)
```
We said earlier that R is a bit like an obedient robot; it will do exactly what you tell it to do.

[HOVER OVER YELLOW TRIANGLE; TWO WARNINGS]

[RUN CODE TO SHOW THAT IT DOESN'T WORK]
[FIX ERROR IN FUNCTION NAME; POINT OUT IT'S NOW BLUE]
[RUN CODE TO SHOW THAT IT STILL DOESN'T WORK]

This error message is quite helpful. If you haven't spotted the problem already, the error message gives you a clue. It tells you that there's no package called `todyverse`, which should be enough for you to realise that you meant 'tidyverse'.

[FIX ERROR IN PACKAGE NAME, EXECUTE LINE; POINT OUT GREEN BAR]

Explain edit/execute loop.

```{r, eval=F}
library(psydata
```

[HOVER OVER RED CROSS]

R Studio tries to be helpful, but doesn't always correctly identify the error. Error messages can be misleading. Here, the suggestion is that we've missed a comma. In actual fact, we've forgotten to close the parentheses to the `library()` function.

### Common errors

* Missing pipe
* Missing assignment
* Missing comma
* %>% not + in ggplot()
* forgot library(tidyverse)
* PSYDATA should be lowercase

### Style

[HOVER OVER BLUE I]

spaces around `=` makes code easier to read.