Sometimes we need to link together multiple steps in our analysis.

For example, if we're working with a big dataset we might want to select only some of the columns, and then filter out some of the rows of data, and the finally calculate descriptive statistics.

We could do this by creating lots of variables, each one saving the results at each intermediate step. This can get confusing, though.

Instead we can use what's known as a 'pipe' — it's another way to link together multiple instructions.

The pipe sends data from one piece of code to another.

The pipe looks like this `%>%`.  

In session 1, you used this code to "pipe" the `mtcars` dataset into `head()`, which shows just the first few rows:

```{r}
mtcars %>% head()
```

You can think of your data as flowing along lengths of pipe, joined by functions which do things to the data, step by
step, until the result you want plops out at the end.

Each `%>%` should be read as the word "then", e.g. "pipe `mtcars` data, *then* `head()` it".

The `>` in the pipe function reminds you of the direction in which your data is flowing (it only works left to right).

It's important to know that the pipe *doesn't store the results* of these steps.

Sometimes that's OK. In our first example we just wanted to look at the first few rows of the `mtcars` data.

But, you will usually want to save the result of a pipeline in a new variable.  
For example, if we wanted to save the first few rows of the `mtcars` data  to a *new* variable we would write:

```{r}
mtcars_head <- mtcars %>% head()
```

Here we combine *assignment* with a *pipeline*.

The result of the pipeline (a `data.frame` containing the first few rows of `mtcars`) is saved to a
new variable called `mtcars_head`.

You can explore your variables using the **Environment** pane. A `data.frame` will have an icon that looks like a
spreadsheet. If you [click on the icon], the `data.frame` is displayed in a new tab in the **Source** pane.

This tab shows you the same information as printing the `data.frame`, such as the number of rows and columns, but it
also provides tools for exploring the data interactively.

* The arrows next to the column names allow you to arrange the rows in ascending or descending order based on the column
values.
* The `Filter` button allows you to specify a value for one or more columns to filter out non-matching rows. For
example, we could display just cars with 4 gears. Click the button again to turn off the filter.
