A scatterplot shows the *relationship* between two variables by plotting their values as points on an x axis (the left-right position) and y-axis (up down).


This chunk creates a scatterplot.

We start with the `mtcars` data.

The `%>%` symbol is special, it's called into a pipe. We'll see more about this in the next session,
but for now just know that it sends the `mtcars` data on to the next line of code — like it's passing
it down a pipe.

```{r}
mtcars %>%
  ggplot(aes(wt, mpg)) +
    geom_point()          
```

[show the code, select the pipe with cursor]


The second line recieves the data. The `ggplot()` function tells us we are going to be making a plot.

The plot is built in two steps.

The first step, `ggplot(aes(wt, mpg))` selects columns in our dataset to use for the x and y axes.

In this case, the x axis is `wt`, which is the weight of the cars

[selects wt in code]

And `mpg` is miles per gallon, or fuel efficiency. This will be the y-axis.

We can see the plot if we put our cursor on the first line and press the shortcut — Ctrl or Cmd + Enter

[run the plot and show]

As you've seen before, if we run code in an Rmarkdown document then the result is shown underneath the chunk.

-------


A useful thing to know is that `ggplot` works by building up plots in multiple layers.

If we run just this part of the code, we can see the plot with just the axes, and no data shown.

[ run just first two lines of code ]

So, we make plots by first defining the axes, and then adding the data points.
Each step is separated by `+` and goes on a new line.
Because RStudio knows this is all part of the same statement it automatically indents the code.



-------------


There's just one final thing to explain:  In the previous code we wrote `x = wt` and `y = mpg`.
This makes things explicit, which is nice, but takes longer to type.

You can also write the plot this way:

```{r}
mtcars %>%
  ggplot(aes(wt, mpg)) +
    geom_point()         
```

R assumes that the first variable is the x axis and the second is the y axis.

[select x axis and y axis in turn when describing]

We use this style in these guides.
