
As we saw before, scatterplots show the relationship between two variables.

Using the `mpg` data, we could plot the size of car engines (`displ`, short for displacement) against their fuel efficiency on
the highway in miles per gallon (`hwy`):

```{r, eval=F}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point()
```


*NOTE: This first plot is not shown here to save space, but is shown in the video.*



Colour can be added to distinguish the points — for example, what kind of car each point represents.

So far, you've used the `aes()` function to define which variable is plotted on the x and y axes.  The `aes()` function is short for 'aesthetics' (what we can see) and *connects columns in our data to visual aspects of a plot*, like the x and y axes.

The `colour=...` option adds to this, and says which column is used to colour the points.
In the `mpg` data, the `drv` column records if a car is front (`f`), rear (`r`) or four (`4`) wheel drive.

We can write `color=drv` to tell R to colour each point, depending on which wheels are driven:

```{r, eval=T}
# colour each point; different colour for each type of transmission
# front, rear and four wheel drive
mpg %>%
  ggplot(aes(displ, hwy, color=drv)) +
  geom_point()
```

This is an example of using a *categorical* variable to colour our points.

Notice that `ggplot()` automatically adds a key to the plot, mapping the colour of the points to the categories in `drv`.
