A scatterplot shows the *relationship* between two variables by plotting their values as points on an x axis (the left-right position) and y-axis (up down).


This code chunk creates a scatterplot. We start with the `fuel` data.

The `%>%` symbol is special, it's called a 'pipe'. We'll explain more later, but for now just know
that it sends the `fuel` data on to the next line of code — like it's passing it down a pipe.

```{r}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()
```

[SHOW THE CODE, SELECT THE PIPE WITH CURSOR]


The second line recieves the data. The `ggplot()` function tells us we are going to be making a plot.

The plot itself is built in two steps. The first step, `ggplot(aes(weight, mpg))` selects columns in our dataset to use for the x and y axes. In this case, the x axis is `weight`, which is the weight of the cars in kg.

[SELECT WEIGHT IN CODE]

And `mpg` is miles per gallon, or fuel efficiency. This will be the y-axis.

We can see the plot if we put our cursor on the first line and press the shortcut — Ctrl or Cmd + Enter

[RUN THE PLOT AND SHOW]

As you've seen before, if we run code in an Rmarkdown document then the result is shown underneath the chunk.



### Building plots in layers


A useful thing to know is that `ggplot` works by building up plots in multiple layers.

If we run just this part of the code, we can see the plot with just the axes, and no data shown.

[RUN JUST THE FIRST TWO LINES OF CODE BY SELECTING AND PRESSING CTRL+ENTER]

So, we make plots by:

1. selecting data
2. defining the axes, and then
3. adding the data points

Each part of the plot is separated by a `+` symbol and goes on a new line.

RStudio is smart and knows all this is part of the sample plot, so automatically indents the code.

### Cutting corners

There's just one final thing to explain:  In the previous code we wrote `x = weight` and `y = mpg`.

This makes things explicit, which is nice, but takes longer to type. You can also write the plot this way:

```{r, eval=F}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()         
```

R assumes that the first variable is the x axis and the second is the y axis.

[SELECT X AXIS AND Y AXIS IN TURN WHEN DESCRIBING]

We use this style in these guides, and you should too.
