A scatterplot shows the *relationship* between two variables by plotting their values as points on an x axis (the left-right position) and y-axis (up down).


This code chunk creates a scatterplot. We start with the `fuel` data.

The `%>%` symbol is special, it's called a 'pipe'. We'll explain more later, but for now just know
that it sends the `fuel` data on to the next line of code — like it's passing it down a pipe.

```{r}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()
```

:::{.direction}

- show the code
- select the pipe with the cursor to highlight

:::


The second line recieves the data. The `ggplot()` part means we're making a plot.

The plot is built in two steps:

The first step is `ggplot(aes(weight, mpg))`. This selects columns in our dataset to use for the x and y axes. In this case, the x axis is `weight`, which is the weight of the cars in kg. And `mpg` is miles per gallon, or fuel efficiency. This will be the y-axis.

:::{.direction}

- select `weight` in the code - highlight it is the x axis
- same for `mpg` and y axis

:::


We can see the plot if we run this statement using the keyboard shortcut — `Ctrl` or `Cmd` (Mac) + `Enter`

:::{.direction}

- run the code and show the resulting plot
- emphasise this is shown below the code chunk when using RMarkdown

:::




### Building plots in layers


A useful thing to know is that `ggplot` works by building up plots in multiple layers.

If we run just this part of the code, we can see the plot with just the axes, and no data shown.


:::{.direction}

- run just the first two lines of code by selecting and pressing `ctrl+enter`
- emphasise the axes are there but no data shown

:::


So, conceptually, we make plots by:

1. *selecting* data
2. *defining* the axes, and then
3. *drawing* the data points

Each part of the plot is separated by a `+` symbol and goes on a new line.

RStudio is smart and knows all this is part of the same statement.
This means it automatically indents the code.



### Cutting corners

There's just one final thing to explain:  In the previous code we wrote `x = weight` and `y = mpg`.

This makes things explicit, which is nice, but takes longer to type. You can also write the plot this way:

```{r, eval=F}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()         
```

R assumes that the first variable is the x axis and the second is the y axis.

:::{.direction}

- select x axis and y axis in turn when describing

:::


We use this style in these guides, and you should too.
