TODO - base video on code examples and summary

-----------------

NOTES


Show

```{r}
mtcars %>% summarise(average_mpg = mean(mpg))
```

and also

```{r}
mtcars %>% summarise(mean(mpg))
```

emphasising that the former is better

Also show replacing with median, sd etc.

This is where we fist encounter the need to give things R-legal names so explain about spaces and special characters
Emphasise everything should be lower case with underscores.

Then show calculating two new cols at once:

mtcars %>% summarise(M = mean(mpg), SD = sd(mpg))

Point out the comma between the two in commentary




### Using `filter` with `summarise`

One powerful thing about R and the `tidyverse` is that you can combine the techniques 
you learn to make more sophisticated analyses.

A simple example of this is using  `filter` with `summarise`. 

For example, if we want to calculate the average life expectancy of only Asian countries
in the development data we can write:

```{r}
development %>% 
  filter(continent=="Asia") %>% 
  summarise(asian_life_expectancy = mean(life_expectancy))
```
