The video introduces:

- the different types of data we may have (e.g. continuous, categorical, text)
- the way R stores these data and
- the way that `ggplot` presents them


These might seem like small details — but having some understanding of what is happening will help later on: you
will be able to control how your data are presented in tables and plots. It will also help you make sure
your data are in the right format for other statistical tests.


### Columns vs variables

:::{.tip}

This is really important. Make sure you are clear on the different meanings of the word 'variable'

:::

The word *variable* gets used in at least 4 different ways in quantitative research, and this confuses a lot of people.
We can't avoid all ambiguity because these different usages are common in the field, but it helps to know
about them in advance:


- *variables in a theoretical model*. That is, things we think exist, and which *cause* other things to happen. An example would be "empathy", or "working memory".

- *variables in a study design*. For example an experimental group allocation, or an attribute of participants like age or gender.

- *variables in a dataset*. Here we mean a column of numbers in a spreadsheet, where the column has a name. This usage overlaps with the previous two, but doesn't have to. For example, we might have a column of numbers recording participants' scores on an empathy questionnaire. But we might also have a column containing reaction times for multiple trials; analysed a certain way these reaction times might tell us something about working memory, but the column of reaction times isn't the same thing as the variable in our theoretical model.

- *variables in R and computer programs*: This type of variable are a *general purpose container* which can store anything. Variables in R *can* contain columns of numbers, but they can also contain whole datasets, or graphs, or the results of statistical tests. **A variable in R often isn't the same thing as an experimental variable or a column in a spreadsheet**.


**In this guide use *variable* to mean either an experimental/theoretical variable or an R container.**
**When we mean a column of data in a dataset we will use the word *column* instead.**



### Types of variable

If we are thinking about variables in a study design the main *types* to consider are:

- *categorical* or *nominal* variables (sometimes also called *factors*): e.g. gender, education status
- *binary* variables (a special category that is either `TRUE` or `FALSE`): e.g. smoker/non-smoker
- *ordinal* variables (categorical responses where the order is important): e.g. a 1-7 response on a Likert-style question
- *interval* or *continuous* variables: e.g. weight in kg, time in milliseconds
- *count* variables (a frequency, measured in whole numbers): e.g. number of social contacts per week
- *text* variables: e.g. free text responses to open-ended questionnaire items



### Types of column

Datasets have multiple columns, each with a unique name.

There are different types of column in R. These data-types are mostly determined by the sort of variable.
But there is some overlap, and the same variable could be stored in different types of column. We sometimes also
need to convert between data types.

There are three data-types you need to know about in R:

- *Factors*, which are used for categorical data (and sometimes ordinal data too)
- *Text*, also called *character* or *string* data.
- *Numeric* columns can called be either *integer* or *double*. Double is computer speak for 'double-precision number', which just means decimals and really big numbers are allowed
- *Logical*, which can be used to store values that can only be either True or False.


The important thing to know is that, although different types of variable are usually stored in the same
type of column in R, *this isn't always the case*. Sometimes data can be stored as the wrong type---e.g.
categorical data can end up stored as continuous--- and in these cases we need to convert from one
format to another.


You can see which data-type is used to store a variable when using the `glimpse` command introduced in
session 1 (see [here](session-1.html#exploring)).

```{r}
iris %>% glimpse
```

In the `glimpse` output you can see the variable names listed on the left, followed by grey text surrounded by angle brackes, e.g.: `<dbl>` which is the abbreviated data-type.

In this built-in dataset, most of the data is numeric (`dbl`), but the `Species` variable is categorical, and stored as a factor (`fct`).


It's possible to convert columns from one type to another, to suit our needs. We'll see more of this later.


| Data/column type | Useful for             | Abbreviations used/subtypes | Often need to convert from |
| ---------------- | ---------------------- | --------------------------- | -------------------------- |
| Numeric          | Continuous, ordinal    | `int`, `dbl`                | Categorical; Text          |
| Factor           | Categorical, ordinal   | `fct`, `ord`                | Text                       |
| Text             | Free text, categorical | `chr`                       | Categorical                |
| Logical          | Binary/boolean         | `lgl`                       | Numeric                    |




### Column types determine the scales on graphs

If we look at the `fuel` data we can see that most of the columns are stored as numeric data (`dbl`):

```{r}
fuel %>% glimpse()
```

This is fine if we want to make a scatter plot. In the example below we have 'miles per gallon' vs the weight of cars in the `fuel` dataset (in `psydata`):

```{r}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()       
```

In this scatterplot both the x and y axes are ***continuous***. 
That is, they are numeric variables, using real numbers.


However, if we want to make a boxplot of the `mpg` column using `cyl` (number of cylinders) as the `x` axis then we have a problem:

```{r}
fuel %>%
  ggplot(aes(cyl, mpg)) +
  geom_boxplot()
```

The warning message says `Continuous x aesthetic -- did you forget aes(group=...)?`. This is a 
clue as to what has gone wrong.

We might have expected to see;

- miles per gallon on the $y$ axis
- three separate boxes, one each for 4, 6 and 8 cylinders

This doesn't work as expected though.

R spots that `cyl` is stored as numeric data, so it creates a continuous scale on the $x$ axis. 
Only one box is drawn at the midpoint of all the values of `cyl`;  because `cyl` ranges from 
4 to 6 the box appears at 6.


We want to use `cyl` as a categorical variable, so we should convert it to a `factor`.
Then our plot will work properly.

We can use the command `factor(cyl)` to tell R that the $x$-axis is a factor:

```{r}
fuel %>%
  ggplot(aes(factor(cyl), mpg)) +
  geom_boxplot()
```

This gives us the boxplot we were expecting. The only change was to replace `cyl` with `factor(cyl)`. 
This tells R to convert the variable `cyl` to a factor, and  `ggplot` can then set the scale of the x axis correctly.
