The video introduces the link different types of data (e.g. continuous, categorical, text),
the way R stores them, and the way that `ggplot` presents them
on the scales of a plot.

These might seem like small details — but having some understanding of these details will help later on.


### Columns vs variables

The word *variable* gets used in at least 4 different ways in quantitative research, and this confuses a lot of people.
We can't always avoid this ambiguity because the usages are so common in the field, but it helps to know about them in advance:


- *variables* in a theoretical model. That is, things we think exist, and which *cause* other things to happen. An example would be "empathy", or "working memory".

- *variables* in a study design. For example an experimental group allocation, or an attribute of participants like age or gender.

- *variables* in a dataset. Here we mean a column of numbers in a spreadsheet, where the column has a name. This usage overlaps with the previous two, but doesn't have to. For example, we might have a column of numbers recording participants' scores on an empathy questionnaire. But we might also have a column containing reaction times for multiple trials; analysed a certain way these reaction times might tell us something about working memory, but the column of reaction times isn't the same thing as the variable in our theoretical model.

- *variables* in R: these are a general purpose *container* which can store anything. Variables *can* contain columns of numbers, but they can also contain whole datasets, or graphs, or the results of statistical tests. A variable in R often isn't the same thing as an experimental variable.


In this guide use *variable* to mean either an experimental/theoretical variable, or an R container. When we mean a column of data in a dataset we will use the word *column* instead.



### Types of variable

If we are thinking about variables in our study designs, the main types are:

- *categorical* or *nominal* variables: these are sometimes also called *factors* in experimental design)
- *binary* variables: these are either true or false, and are a special kind of categorical data
- *interval* or *continuous* variables: e.g. heights, weights, or reaction times
- *ordinal* variables: e.g. a response to a Likert-style question, on a 1-7 scale
- *count* variables: how often something has occurred, measured in whole numbers greater than or equal to zero



### Types of column

Datasets have multiple columns, each with a unique name.

There are different types of column in R. These data-types are mostly determined by the sort of variable.
But there is some overlap, and the same variable could be stored in different types of column. We sometimes also
need to convert between data types.

There are three data-types you need to know about in R:

- *Factors*, which are used for categorical data.
- *Text*, also called *character* or *string* data.
- *Numeric* columns can called be either *integer* or *double*. Double is computer speak for 'double-precision number', which just means decimals and really big numbers are allowed
- *Logical*, which can be used to store values that can only be either True or False.


You can see which data-type is used to store a variable when using the `glimpse` command you saw in session 1 (e.g. [here](session-1.html#exploring)).

```{r}
iris %>% glimpse
```

In the `glimpse` output you can see the variable names listed on the left, followed by grey text surrounded by angle brackes, e.g.: `<dbl>` which is the abbreviated data-type.

In this built-in dataset, most of the data is numeric (`dbl`), but the `Species` variable is categorical, and stored as a factor (`fct`).


It's possible to convert columns from one type to another, to suit our needs, and we'll see more of this later.


| Data/column type | Useful for             | Abbreviations used/subtypes | Often need to convert from | 
| ---------------- | ---------------------- | --------------------------- | -------------------------- |
| Numeric          | Continuous, ordinal    | `int`, `dbl`                | Categorical; Text          |
| Factor           | Categorical, ordinal   | `fct`, `ord`                | Text                       |
| Text             | Free text, categorical | `chr`                       | Categorical                |
| Logical          | Binary/boolean         | `lgl`                       | Numeric                    |




### Column types and scales on graphs

If we look at the mtcars data we can see that all the columns are stored as numeric data (`dbl`):

```{r}
mtcars %>% glimpse()
```

This is fine if we want to make a scatter plot (here of 'miles per gallon' vs the weight of the car):

```{r}
mtcars %>%
  ggplot(aes(wt, mpg)) +
    geom_point()       
```

In this plot both the x and y axes are ***continuous***. That is, they are numeric variables, using real numbers.


However, if we want to make a boxplot of the `mpg` column using `am` as the `x` axis then we have a problem:

```{r}
mtcars %>%
  ggplot(aes(am, mpg)) +
  geom_boxplot()
```

We might have expected to see;

- miles per gallon on the y axis
- two separate boxes, one for automatic cars and another for manual.

This doesn't work as expected though. 

R spots that `am` is stored as numeric data, so it creates a continuous scale on the x axis. Only one box is drawn at the midpoint of all the values of `am`;  because `am` ranges from 0 to 1 the box appears at 0.5.



We want to use `am` as a categorical variable, so we should convert it to a factor.
Then our plot will work properly.


We can use the command `factor(am)` to tell R that the x-axis is a factor:

```{r}
mtcars %>%
  ggplot(aes(factor(am), mpg)) +
  geom_boxplot()
```

This gives us the boxplot we were expecting. The only change was to replace `am` with `factor(am)`. This tells R to convert the variable `am` to a factor, and  `ggplot` can guess the type of x axis correctly.
