R is a computer language for data analysis and visualisation.

RStudio is a user interface to R; it helps you organise your work.

R is a text-based language. You interact with it by typing commands and running (or executing) them.


R can do everything from simple arithmetic and plotting to complex data analysis.

For example, you can do simple arithmetic like

[`2 * 221`]

We could generate some random numbers with a [normal distribution](https://www.khanacademy.org/math/statistics-probability/modeling-distributions-of-data/normal-distributions-library/a/normal-distributions-review)

```{r}
rnorm(10, 0,1)
```


And we could plot random numbers using a histogram

```{r}
hist(rnorm(100, 0,1))
```

------------

You should think of R as a robot.

The robot is extremely fast, powerful and tireless; but it’s also literal-minded, and
won’t think for itself or take the initiative. You need to tell it *exactly* what to do, by providing very precise
instructions.

The advantage of writing detailed instructions is that you have a detailed, reproducible
version of all your analyses.

Reproducibility is a key topic in psychology and other natural sciences — learning R (or something like it)
is an important skill for new psychologists.
