---
title: "LifesaveR: Session 1"
output: html_document
---

# lifesaveR: Workbook 1

This is an RMarkdown document. We will use a document like this as a workbook for 
each of the sessions in the LifesaveR course.

What to do: 

- Use this document to complete and record your answers to the exercises in each session.
- Edit the code and text below to save your work



## An example plot

The following chunk creates a scatter-plot of a built-in data set that contains information about diamonds. By the end of
the second session, you will have learnt how to create similar plots.

```{r}
library(tidyverse)
diamonds %>%
  ggplot(aes(carat, price)) +
  geom_point()
```

# Working interactively in R Markdown {#interactive}

## Running R code within a chunk

### Exercise 1

1. Locate the first chunk in this workbook.
1. Place your cursor (anywhere) on the line that says `library(tidyverse)`.
1. Run the code.

### Exercise 2

Position your cursor on the line that says `diamonds` and run the code.

```{r}
diamonds %>%
  ggplot(aes(carat, price)) +
  geom_point()
```

### Exercise 3

1. Select (highlight) the word `diamonds` in the chunk above.
1. Run the code.

## Inserting a chunk

### Exercise 4

1. Create a chunk below these instructions
1. Inside the chunk, write a line of code which adds together the numbers 9, 4, 55 and 2
1. Run the the line of code you have written

# Built-in datasets

# Check your knowledge

## How do you run part of a line of R code?

[Write your answers directly into the workbook here and save it for later]

## Which library do you need to load in your first R Markdown chunk?

## What is `gapminder`?

## Which function would you use to explore the first few rows of a dataset?

## What is the 5th column in `gapminder`?

## Explain what `glimpse` does.

## Which function makes a plot?

## Which function defines the axes on a plot?

## Which function makes a boxplot?

## How do you use markdown to make a level 3 heading?