---
title: 'Working with groups'
author: 'Ben Whalley, Paul Sharpe, Sonja Heintz'
# date: "April 2021"
bibliography: [references.bib]
csl: apa.csl
#biblio-style: apa6
link-citations: yes
output:
  webex::html_clean:
    highlight: tango
---

```{r, include=F, echo=F}
source('_first_chunk.R')
```


```{html, child="bs-tab-fix.html"}
```


```{css, child="video.css"}
```


```{r, fig.height=2, fig.width=12, echo=F, message=F, warning=F}
development %>%
  ggplot(aes(fct_reorder(country, as.numeric(continent), median, .desc=TRUE), life_expectancy, colour=continent)) +
  geom_boxplot() +
  xlab("") + ylab("") +theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +guides(colour=F) + theme(
          panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = unit(c(0, 0, 0, 0), "null"),
    panel.margin = unit(c(0, 0, 0, 0), "null"))

```


# Overview

As researchers, we often collect data from two or more groups.

Group allocations are *categorical* variables. They are stored in a special way by R, which makes it easier to display
them on graphs and use them in analyses.

In this workshop we learn to how to use categorical variables in boxplots, to add colour to scatterplots, and
to make tables of descriptive statistics which you will need in your coursework and project reports.



### Techniques covered

- [Making boxplots](#boxplots)
- [Adding colour to a plot](#colour)
- [Using different types of data and visual scales](#datatypes)
- [Using `group_by()` to make a summary table and compare groups](#groupby)


# Making boxplots {#boxplots}

`r video("2FXgGkfXtXw")`
`r make_tabs('makingboxplotstabs')`

:::{.tab-content}

:::{#makingboxplotstabs-summary .tab-pane .active}

- Boxplots are useful for comparing categorised data
- Use `aes()` with `ggplot()` to choose a categorical column as the x-axis
- `geom_boxplot()` draws boxes and whiskers for each category
- The box describes the [*interquartile range*](https://www.bbc.co.uk/bitesize/guides/zx7rk7h/revision/2)
- The midpoint is the *median*
- Individual points show *outliers* in the data

:::

:::{#makingboxplotstabs-code .tab-pane .fade}
```{r}
iris %>%
  ggplot(aes(x=Species, y=Petal.Length)) +  # specify dataset columns 
  geom_boxplot()                            # add the boxes
```
:::


:::{#makingboxplotstabs-explanation .tab-pane .fade}
Boxplots are useful for visualising differences between categories or groups.

Using `ggplot()`, making a boxplot is similar to a scatterplot.

If we haven't already done it, we'd need to load the `tidyverse`:

```{r, echo=T, warning=F, message=F}
library(tidyverse)
```

----------------------------------------------------------------------------------------------------

To make the plot, we first choose the data we want to plot and add a pipe to the end of the line to
send this data to the next command:

::: direction
-   From the code below select only first line
-   Also show selecting/executing just the name of the dataset and how this displays the data
-   Emphaise the importance of breaking code down and running sub-steps separately
:::

Next we use `ggplot()` and `aes(x = ..., y = ...)` to select the columns we want to use in the data.
`aes` is short for *aesthetics* which means 'something you can see'. So the `aes()` function defines
what will be able to see on the plot --- the axes, and other features like colour.

This time we have chosen `Species` for the x-axis, because it is a *categorical* variable.

```{r}
iris %>%
  ggplot(aes(x = Species, y = Petal.Length))
```

::: direction
-   Demonstrate running code above
-   Also show using autocomplete in RStudio when writing the code above
:::

You'll notice that I used the RStudio autocomplete feature to write function and column names from
our dataset. This makes things much easier --- especially for column names. If you type column names
by hand it's easy to make spelling errors or typos, and this leads to errors in R.

If we run the code so far, `ggplot()` draws the plot axes, but doesn't add the data yet:

::: direction
Run the code so far
:::

So far we have said *what* we want to show on our plot, but not *how* it should be shown.

To finish, we add `geom_boxplot()` which actually draws the boxes:

```{r}
iris %>%
  ggplot(aes(x = Species, y = Petal.Length)) +
  geom_boxplot()
```

Note that I used a `+` symbol to add the boxplot layer to our graph.

So, now we have a boxplot, with one box drawn for each value of `Species`.

### Interpreting boxplots

In a boxplot:

-   The thick line across the middle of the box is the *median* or midpoint of the data

-   The height of the box indicates the [*interquartile
    range*](https://www.bbc.co.uk/bitesize/guides/zx7rk7h/revision/2) (IQR). This means the box
    contains 50% of the datapoints. A wider IQR indicates greater variation (spread) in a dataset.

-   The meaning of the "whiskers" (the lines above the top, and below the bottom of the box) varies
    a little bit depending on the software you use --- but they always show the range/spread of the
    data. The default in `ggplot()` are whiskers with lengths covering data points no more, or less
    than 1.5 times the IQR.

-   Any data point outside the range of the whiskers is described as an 'outlying point' or
    'outlier'. Each outlier is plotted individually as a dot.

:::

:::



:::{.exercise}
**Exercise 1**

1. Open `session-3.rmd` using the **Files** pane. This is the workbook you will be using in this session.
1. Use the `iris` dataset to create a boxplot with `Species` on the x-axis and `Sepal.Width` on the y-axis (sepals are the leaves that encase an iris flower).

Your boxplot should look like this:

```{r, echo=FALSE}
iris %>%
  ggplot(aes(Species, Sepal.Width)) +
  geom_boxplot()
```
:::

# Adding colour to a plot {#colour}


`r video("wzjGY-Yn2g8")`
`r make_tabs('scatterplotcolortabs')`

:::{.tab-content}

:::{#scatterplotcolortabs-summary .tab-pane .active}

- The points of a scatterplot can be coloured based on a third column of data
- We can colour points using categorical ***or*** continuous data
- The ` aes(..., colour=column_name)` option selects the column to use (change `column_name` to the name of *your* column)

:::

:::{#scatterplotcolortabs-code .tab-pane .fade}
```{r}
# scatterplot of engine size (displ) against highway fuel efficiency (hwy)
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point()

# colour each point; different colour for each type of transmission
# front, rear and four wheel drive
mpg %>%
  ggplot(aes(displ, hwy, colour=drv)) +
  geom_point()
```
:::


:::{#scatterplotcolortabs-explanation .tab-pane .fade}
As we saw before, scatterplots show the relationship between two variables.

Using the `mpg` data, we could plot the size of car engines (`displ`, short for displacement)
against their fuel efficiency on the highway in miles per gallon (`hwy`):

```{r, eval=F}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point()
```

::: direction
-   Run code to generate plot
:::

*NOTE: This first plot is not shown here to save space, but is shown in the video.*

We can add colour to distinguish the points in this plot --- for example, what kind of car each
point represents.

So far, you've used the `aes()` function to define which variable is plotted on the x and y axes.
The `aes()` function is short for 'aesthetics' (what we can see) and *connects columns in our data
to visual aspects of a plot*, like the x and y axes.

The `colour=...` option adds to this, and says which column is used to colour the points.

In the `mpg` data, the `drv` column contains one of three values. The value categorises cars based
on which wheels drive the engine. An `f` in `drv` means a front wheel drive car, `r` means rear
wheel drive, and `4` means four wheel drive.

We can write `colour=drv` to tell R to colour each point, depending on which wheels are driven:

::: direction
1.  Add `colour=drv`
2.  Run code to generate plot
:::

```{r, eval=T}
# colour each point; different colour for each type of transmission
# front, rear and four wheel drive
mpg %>%
  ggplot(aes(displ, hwy, colour=drv)) +
  geom_point()
```

Notice that `ggplot()` automatically adds a key to the plot, mapping the colour of the points to the
categories in `drv`.

This is an example of using a *categorical* variable to colour our points.

:::

:::






:::{.exercise}
**Exercise 2**

1. Use the `development` data from the `psydata` package (run `library(psydata)` first if you haven't already).
1. Create a scatterplot with `life_expectancy` on the x-axis (along the bottom) and `gdp_per_capita`
   on the y-axis.
1. Add colour to this scatterplot: make each `continent` a different colour.
1. Run the chunk of code.

Your plot should look like this:

```{r ex10, echo=FALSE}
library(psydata)
development %>%
  ggplot(aes(life_expectancy, gdp_per_capita, colour = continent)) +
  geom_point()
```
:::




# Using different types of data and visual scales{#datatypes}


`r video("RrIRNof4mzM")`
`r make_tabs('typesscalestabs')`

:::{.tab-content}

:::{#typesscalestabs-summary .tab-pane .active}

- Continuous, categorical and text data are all common in psychology
- Internally, R stores columns of data using different *data types*.
- R uses data types as a clue to set *defaults* (e.g. for graphs)
- Sometimes we need to convert between types
- For example, sometimes categorical data is (wrongly) stored as a numeric column

:::

:::{#typesscalestabs-code .tab-pane .fade}
```{r}
# always load the libraries we need first
library(tidyverse)
library(psydata)

# see a list of columns in the dataset, and their types
iris %>% glimpse

# make a scatter plot with two continous axes
# both wt and mpg are numeric variables
# so this works well
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()   

# try and make a boxplot with `am` as the x-axis
# but because `cyl` is stored as a numeric variable
# (not categorical) the scale of the x-axis is wrong
fuel %>%
  ggplot(aes(cyl, mpg)) +
  geom_boxplot()

# re-draw the plot, but converting `cyl` to a factor/categorical
# variable first. Now the x-axis looks correct
fuel %>%
  ggplot(aes(factor(cyl), mpg)) +
  geom_boxplot()
```
:::


:::{#typesscalestabs-explanation .tab-pane .fade}
```{=html}
<!--
TODO - have a ppt of the first explanatory part of this instead of a blank screen
 -->
```
Ok this is a slightly longer video. In the first bit we introduce some background to the term
variable, and the way in which R stores different types of data. In the final part we see why this
is important when we're making plots (and running other analyses).

I'm especially going to talk about:

-   the different types of variable we may have in our study (e.g. continuous, categorical, text)
-   the way R stores these different data (e.g. as numeric, factor or character data) and
-   the way that `ggplot()` presents these different data types

These might seem like small details --- but having some understanding of what is happening will help
later on: you will be able to control how your data are presented in tables and plots. It will also
help you make sure your data are in the right format for other statistical tests.

### Columns vs variables

::: tip
This is really important. Make sure you are clear on the different meanings of the word 'variable'
:::

The word *variable* gets used in at least 4 different ways in quantitative research, and this
confuses a lot of people.

We can't avoid all ambiguity because these different usages are common in the field, but it helps to
know about them in advance. The main usages are:

-   *variables in a theoretical model*. That is, things or constructs we think exist, and which
    *cause* other things to happen. An example might be "empathy", or "working memory". These are
    psychological constructs which many theories include, and those theories say these constructs
    play a causal role in how we behave or perform on certain tasks.

-   *variables in a study design*. An example here might an experimental group allocation, or an
    attribute of participants like their age or gender.

-   *variables in a dataset*. Sometimes people want to refer to a column of numbers in a dataset
    (e.g. a spreadsheet), where the column has a name, and use the term variable again. This usage
    overlaps with the previous two, but it doesn't *have* to. For example, we might have a column of
    numbers recording the date and time participants' completed an empathy questionnaire. Analysed a
    certain way these timestamps *might* tell us something useful, but the column of timestamps
    isn't a variable in our theoretical model, and they might not be used as a variable in our
    experiment.

[ Finally, we have ]

-   *variables in R and computer programs*: This type of variable is a *general purpose container*
    which can store anything. Variables in R *can* contain columns of numbers, but they can also
    contain whole datasets, or graphs, or the results of statistical tests. **A variable in R often
    isn't the same thing as an experimental variable or a column in a dataset**.

**In this short course use *variable* to mean either an experimental/theoretical variable or an R
container.**

**When we mean a column of data in a dataset we will use the word *column* instead.**

### Types of variable

OK so let's focus for a minute on *variables in a study design*.

There are a few main *types* of these variable to consider. You might be familiar with some of these
terms already:

-   *categorical* or *nominal* variables (sometimes also called *factors*): e.g. gender, education
    status

-   *binary* variables (a special category that is either `TRUE` or `FALSE`): e.g. smoker/non-smoker

-   *ordinal* variables (categorical responses where the order is important): e.g. a 1-7 response on
    a Likert-style question

-   *interval* or *continuous* variables: e.g. weight in kg, time in milliseconds

-   *count* variables (a frequency, measured in whole numbers): e.g. the number of social contacts
    per week

-   *text* variables: e.g. free text responses to open-ended questionnaire items

### Types of column

OK, so those are the types of variable.

We can also think about the types of column that our dataset stores.

Datasets have multiple columns, each with a unique name.

In R, every column has a *data type*. These data types are normally determined by the type of
variable. But there is some overlap, and different data types could be used to store the same
variable.

We sometimes need to convert between data types.

There are four data types you need to know about in R: factors, text, numeric and logical columns.

-   *Factors*, which are used for categorical data (and sometimes ordinal data too)
-   *Text*, also called *character* or *string* data.
-   *Numeric* columns can called be either *integer* or *double*. Double is computer speak for
    'double-precision number', which just means decimals and really big numbers are allowed
-   *Logical*, which can be used to store values that can only be either True or False.

The important thing to know is that--- although different types of variable are usually stored in
different types of column in R--- *this isn't always the case*.

Sometimes data can be stored as the wrong type---e.g. categorical data could end up stored as
numeric data --- and in these cases we need to convert from one format to another.

You can see which data type is used to store a variable when using the `glimpse()` command
introduced in Session 1 (see [here](session-1.html#exploring)).

```{r}
iris %>% glimpse()
```

In the output you can see the variable names listed on the left, followed by text in angle brackets,
e.g.: `<dbl>` which is the abbreviated name of data type.

In this built-in dataset, most of the data is numeric (`dbl`), but the `Species` variable is
categorical, and stored as a factor (`fct`).

It's possible to convert columns from one type to another, to suit our needs. We'll see more of this
later.

| Data/column type | Useful for           | Abbreviations used/subtypes | Often need to convert from |
|------------------|----------------------|-----------------------------|----------------------------|
| Factor           | categorical, ordinal | `fct`, `ord`                | text                       |
| Text             | text, categorical    | `chr`                       | categorical                |
| Numeric          | continuous, ordinal  | `int`, `dbl`                | categorical; text          |
| Logical          | binary (boolean)     | `lgl`                       | numeric                    |

### Column types determine the scales on graphs

If we look at the `fuel` data (from `psydata`) we can see that most of the columns are stored as
numeric data (`dbl`):

```{r}
fuel %>% glimpse()
```

This is fine if we want to make a scatter plot.

In the example below we have 'miles per gallon' vs the weight of cars in the `fuel` dataset (in
`psydata`):

```{r}
fuel %>%
  ggplot(aes(weight, mpg)) +
    geom_point()       
```

In this example of a scatterplot both the x and y axes are ***continuous***. That is, we expect them
to be real numbers, and they are stored as numeric columns.

However, if we want to make a boxplot of the `mpg` column using `cyl` (number of cylinders) as the
`x` axis then we have a problem:

```{r}
fuel %>%
  ggplot(aes(cyl, mpg)) +
  geom_boxplot()
```

The warning message says `Continuous x aesthetic -- did you forget aes(group=...)?`. This is a clue
as to what has gone wrong.

We might have expected to see;

-   miles per gallon on the $y$ axis
-   three separate boxes, one each for 4, 6 and 8 cylinders

It didn't work as expected though.

R spots that `cyl` is *stored* as numeric data, and because it doesn't know better it creates a
\_continuous scale) on the $x$ axis.

Only one box is drawn at the midpoint of all the values of `cyl`; because `cyl` ranges from 4 to 6
the box appears at 6.

Although R often has good defaults, it isn't smart --- if it doesn't behave as we expect we need to
be more specific.

Here we want to use `cyl` as a categorical variable, so we need to give R precise instructions. We
should convert `cyl` to a `factor` explictly. Then our plot will work properly.

We can use the command `factor(cyl)` to tell R that the $x$-axis is a factor:

```{r}
fuel %>%
  ggplot(aes(factor(cyl), mpg)) +
  geom_boxplot()
```

This gives us the boxplot we were expecting.

The only change was to replace `cyl` with `factor(cyl)`. This tells R to convert the variable `cyl`
to a factor, and `ggplot()` can then set the scale of the x-axis correctly.

:::

:::




:::{.exercise}
**Exercise 3**

Work with a friend: Describe the 4 ways in which quantitative researchers might use the word 'variable'?

(If you need to look these up from the video or text above then try testing yourself again after completing other exercises).

:::

:::{.exercise}
**Exercise 4**

Use `glimpse()` to check the data types of the `mpg` and the `diamonds` datasets.

- The data type of the 4th variable in the `mpg` data is `r mcq(c("double", "factor", "character", answer=typeof(mpg[[4]])))`
- The data type of the 5th variable in the `diamonds` data is `r mcq(c("date","character", "factor", answer=typeof(diamonds[[5]])))`

:::


:::{.exercise}
**Exercise 5**

Use the `fuel` dataset to make a boxplot showing miles per gallon on the y-axis, and number of gears (`gear`) on the x-axis. Your plot should look like this:

```{r, echo=FALSE}
fuel %>%
  ggplot(aes(factor(gear), mpg)) +
  geom_boxplot()
```

Remember that you will need to *change* the type of the `gear` variable for this to work.

:::





# Using `group_by()` to make a summary table and compare groups{#groupby}

`r video("dBfkA3-j3dU")`
`r make_tabs('groupbytabs')`

:::{.tab-content}

:::{#groupbytabs-summary .tab-pane .active}

- Datasets often contain categorical variables
- We often want to compare statistics (like averages) *between* categories
- The `group_by()` function is a quick way to combine filtering and summarising
- `group_by()` creates a *grouped* `data.frame`
- Adding `group_by()` to a pipeline runs the subsequent steps *once for each group*.
- The result is always a new `data.frame`

:::

:::{#groupbytabs-code .tab-pane .fade}
```{r}
# check the columns in the funimagery dataset
# results of an RCT of functional imagery training for weight loss
# conducted at the University of Plymouth
funimagery %>% glimpse

# boxplot to compare the intervention groups
funimagery %>%
  ggplot(aes(intervention, weight_lost_end_trt)) +
  geom_boxplot()


# calculate mean weight loss in each group in a laborious way
funimagery %>%
  filter(intervention=="MI") %>%
  summarise(mean(weight_lost_end_trt))

funimagery %>%
  filter(intervention=="FIT") %>%
  summarise(mean(weight_lost_end_trt))


# use group_by to split the data and summarise each group
funimagery %>%
  group_by(intervention) %>%
  summarise(mean(weight_lost_end_trt))

# store the result (also a data.frame) in a new variable
average_weight_loss <- funimagery %>%
                          group_by(intervention) %>%
                          summarise(mean(weight_lost_end_trt))


# example of grouping by two columns at once
funimagery %>%
  group_by(gender, intervention) %>%
  summarise(mean(weight_lost_end_trt))



# calculate the mean and SD in one go
funimagery %>%
  group_by(intervention) %>%
  summarise(
    mean(weight_lost_end_trt),
    sd(weight_lost_end_trt)
  )


# how to give your new summary columns a name
# (this is good practice)
funimagery %>%
  group_by(intervention) %>%
  summarise(
    mean_weight_lost_end_trt = mean(weight_lost_end_trt),
    sd_weight_lost_end_trt = sd(weight_lost_end_trt)
  )
```
:::


:::{#groupbytabs-explanation .tab-pane .fade}

In this video we'll use the `funimagery` dataset, which is part of the `psydata` package. This
contains data from a randomised controlled trial run in Plymouth [@solbrig2019functional].

The study compared two treatments for weight loss: Functional Imagery Training (FIT) and
Motivational Interviewing (MI).

We can see the columns in the dataset using `glimpse()`:

::: direction
-   Run code
:::

```{r}
funimagery %>% glimpse
```

Weight was measured three times, and these measurements are in the columns called `kg1`, `kg2` and
`kg3`.

-   The first measurment (`kg1`) was taken at *baseline*, when participants joined the study
-   The second measurement (`kg2`) was taken at the end of treatment, which was 6 months after
    baseline
-   The final measurement (`kg3`) was made 12 months after baseline

The `weight_lost_end_trt` column is the *difference* between `kg1` and `kg2` -- i.e. how much weight
the participant lost after their treatment.

We have already seen a that a boxplot can compare scores between categories:

::: direction
-   Generate plot
:::

```{r}
funimagery %>%
  ggplot(aes(intervention, weight_lost_end_trt)) +
  geom_boxplot()
```

This plot is helpful, and makes the differences really clear. But what if we want the actual numbers
in a table, or to report in the text of our paper?

### Using `filter()` and `summarise()`

One option would be to filter our data first and then summarise each subset in turn:

::: direction
-   Run code
:::

```{r}
funimagery %>%
  filter(intervention=="MI") %>%
  summarise(mean(weight_lost_end_trt))
```

This output shows us the mean for the MI group. We then have to repeat that code for the FIT group:

::: direction
-   Run code
:::

```{r}
funimagery %>%
  filter(intervention=="FIT") %>%
  summarise(mean(weight_lost_end_trt))
```

This is a bit repetitive, and would worse if there were lots of categories. Imagine doing something
similar for each continent in the `development` data, for example.

### Using `group_by()`

Instead of using `filter()`, we can use `group_by()` to split our dataset into multiple groups,
summarising each one separately:

::: direction
-   Run code
:::

```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean(weight_lost_end_trt))
```

Using `group_by()` before `summarise()` splits up the data, in thise case depending on which
`intervention` the participant received.

::: direction
-   Highlight `group_by()`
:::

The `summarise()` function calculates the mean for each group separately.

::: direction
-   Highlight `summarise()`
:::

The result is a new `data.frame` with two columns: The name of the `intervention`, and the average
weight lost. We could save this in a new variable for use later, if we liked:

```{r}
average_weight_losses <- funimagery %>%
                          group_by(intervention) %>%
                          summarise(mean(weight_lost_end_trt))
```

You can see the new variable in the **Environment**, so it's stored for later.

### Nested groups

Sometimes we might want to group by more than one variable at once.

Imagine we wanted to see if men and women responded differently to the two treatments?

With `group_by()` this is really easy: just add the name of another column to group on (`gender`),
separated with a comma:

```{r}
funimagery %>%
  group_by(gender, intervention) %>%
  summarise(mean(weight_lost_end_trt))
```

It looks like women lost a bit more weight on average

::: direction
-   Highlight `f` rows
:::

but both men and women lost more weight after FIT relative to MI.

::: direction
-   Highlight `MI` and `FIT` rows for `f`, then `m`
:::

### Multiple statistics

With `summarise()` we can calculate multiple statistics at once for each of the groups:

::: direction
-   Run code
:::

```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean(weight_lost_end_trt), sd(weight_lost_end_trt))
```

Here, we've calculated both the mean and standard deviation for each intervention.

### Give your new variables a name

R gives the new columns a name based on the function we use to `summarise()` them.

For example, if we use `mean()` on the `weight_lost_end_trt` variable then the new column is called
`mean(weight_lost_end_trt)`.

::: direction
-   Rerun code above and highlight where the name of the variable comes from
:::

Unfortunately, these new variable names contain parentheses and (sometimes) spaces which makes them
harder to work with. It's much better to give the new column a specific name like this:

```{r}
funimagery %>%
  group_by(intervention) %>%
  summarise(mean_weight_lost_end_trt = mean(weight_lost_end_trt), sd_weight_lost_end_trt = sd(weight_lost_end_trt))
```

Remember, the column names you use shouldn't include spaces or other 'special' characters.

:::

:::






:::{.exercise}

**Exercise 6**

Replace the `?` in the code below to to calculate the `median` weight lost for men and women undergoing FIT and MI in the
`funimagery` dataset.

```{r, eval=F}
? %>%
  group_by(?, ?) %>%
  summarise( ?(weight_lost_end_trt) )
```

These are the correct numbers to check your work against:

```{r, echo=F, include=T}
funimagery %>%
  group_by(Gender=gender, Intervention=intervention) %>%
  summarise(`Median weight loss` =  sprintf("%.1f", median(weight_lost_end_trt) )) %>%
  mutate(Gender=factor(Gender, levels=c("m", "f"), labels=c("Male", "Female"))) %>%
  pander()
```


:::


:::{.exercise}

**Exercise 7**

Use `group_by()` and `summarise()` with the built-in `iris` dataset to calculate the mean `Sepal.Length` for each `Species` of flower.

These are the correct numbers to check your work against:

```{r, echo=F, include=T}
iris %>%
  group_by(Species) %>%
  summarise(`Mean sepal length` =  sprintf("%.1f", mean(Sepal.Length) )) %>%
  pander()
```
:::


:::{.exercise}

**Exercise 8**

The built-in dataset `chickwts` contains weights of chicks (in grams) fed on different diets. Use `group_by()` and
`summarise()` to calculate the mean and standard deviation chick weights for each type of feed.

```{r, echo=FALSE, eval=FALSE}
chickwts %>%
  group_by(feed) %>%
  summarise(mean(weight), sd(weight))
```

The mean weight of chicks fed on linseed was `r fitb('218.75', tol=1)`g.

The standard deviation of chicks fed on sunflower was `r fitb('48.84', tol=1)`g.

:::



# Check your knowledge

Write an answer to each of these questions in the `Check your knowledge` section of your workbook. The answers will be revealed in Session 4.

1. Which functions are needed to make a boxplot?
1. What is the difference between a `dbl` and a `fct` or `ord`?
1. Which data types are used for continuous variables?
1. Give an example of why the difference between `dbl` and `fct` matters when making a plot (include code examples for this if you can)
1. How can you convert a variable from a `dbl` to a `fct`?
1. How could you calculate the mean for one level of a factor?
1. How would you calculate the mean for all levels of a factor (e.g. for continent in the `development` dataset)?
1. How would you calculate means and standard deviations for all combinations of levels in two factors?

# Extension exercises


```{r, child="extension-exercises-premable.rmd"}
```




:::{.exercise}

**Extension exercise 1**

Make a scatterplot of the `diamonds` data. Show `carat` on the x-axis, `price` on the y-axis, and the `clarity` of the
diamonds in colour. Try to produce your plot _before_ comparing it against the answer using the button below.

`r hide('show answer')`
```{r, echo=FALSE}
diamonds %>%
  ggplot(aes(carat, price, colour = clarity)) +
  geom_point()
```
`r unhide()`


Interpret your plot: Which category of diamond `clarity` has the steepest rise in price as
size (carat) increases?

:::

:::{.exercise}

**Extension exercise 2**

Make a scatterplot of the `fuel` dataset. Show `mpg` on the y-axis, engine size on the x-axis, and use type of transmission (auto/manual) to colour the points. Try to produce your plot before comparing it against the answer using the button below.

`r hide('show answer')`
```{r, echo=FALSE}
fuel %>%
  ggplot(aes(engine_size, mpg, colour = automatic)) +
  geom_point()
```
`r unhide()`

Interpret your plot: Which type of car (auto or manal) sees the strongest relationship between engine size and fuel economy?
:::

:::{.exercise}

**Extension exercise 3**

Using the `development` dataset, make a boxplot showing life expectancy by continent for years greater than 1999. (Hint: use `filter()`, `ggplot()` and `geom_boxplot()`.)

The plot should look like this:

```{r, echo=FALSE}
development %>%
  filter(year > 1999) %>%
  ggplot(aes(continent, life_expectancy)) +
  geom_boxplot()
```
:::

:::{.exercise}

**Extension exercise 4**

Try to recreate the plot below using the `mtcars` built-in dataset. Remember to use `factor()` for columns which require conversion to `fct`.

```{r, echo=F}
mtcars %>%
  ggplot(aes(wt, mpg, colour=factor(gear))) +
  geom_point()
```
:::

:::{.exercise}

**Extension exercise 5**

Try to recreate the table below using the `painmusic` dataset from `psydata`.

```{r, echo=FALSE}
painmusic %>%
  group_by(liked = factor(liked), familiar = factor(familiar)) %>%
  summarise(mean_pain_no_music = mean(pain.no.music), sd_pain_no_music = sd(pain.no.music))
```

:::

# References
