---
title: "Example data supplement"
output: html_document
authors: "Ben Whalley"
date: '2022-09-06'
---





### Overview

This document provides the analysis to support for Whalley et al, 2022: *"Boring paper on car fuel economy"*.

Each statement in that article requiring statistical analysis is copied below, and
followed by the necessary R code and output.

This supplement is NOT simply all the code used when analysing the data; it 
is an edited and selected subset of that. The analyses here are organised so that
they follow the logic of the paper, making it easier for readers to see where 
individual numbers in the text came from.



### Analyses

```{r}
# this tidies up the output somewhat
knitr::opts_chunk$set(message = F, warning=F)

library(tidyverse)
library(BayesFactor)
library(psydata)
library(pander)
```



```{r}
#model <- modelA
require(broom)

regression_table <- function(model, as_text=TRUE){
  x <- model %>% broom::tidy(conf.int=T) 
  
  z <- model %>% arm::standardize() %>% 
    broom::tidy()
  
  znum <- z %>% 
    mutate(term = str_replace_all(term, "z\\.", "")) %>% 
    transmute(term=term, β = estimate) %>% 
    left_join(x, ., by="term") %>% 
    select(Term=term, Estimate=estimate, `Lower`=conf.low, Upper=conf.high, β, p=p.value)
  
  r2 <- model %>% glance() %>% pull(adj.r.squared)
  znumr2 <- znum %>% 
    bind_rows(., tibble(Term="Adjusted R sq.", Estimate=r2)) 
  
  if(as_text){
    znumr2 %>% 
      mutate(`95% CI` = ifelse(!is.na(Lower), 
        glue::glue("[{sprintf('%.3f', Lower)}, {sprintf('%.3f', Upper)}]"), ""), 
        .after="Estimate") %>% 
      mutate_at(c("Estimate"), ~{ifelse(is.na(.x), "", sprintf("%.3f", .x) )}) %>% 
      mutate_if(is.numeric, ~{ifelse(is.na(.x), "", sprintf("%.2f", .x) )}) %>% 
      mutate_at(c("p"), ~ifelse(is.na(.x)|.x=="", "", apastats::round.p(.x))) %>% 
      select(-Lower, -Upper)
      
  } else {
    znumr2 %>% mutate_if(is.numeric, ~round(.x, 3))
  }
  
}
```



### Specific results from the article

Data were obtained for 32 cars...


```{r}
# note we are using data from psydata package
# normally this would be the point at which you read in a csv or other data file
fuel %>% glimpse
```




Car weights ranged from ...

```{r}
fuel %>% 
  summarise_at(c("mpg", "cyl", "weight"), funs(range, mean, median, sd)) %>% 
  pivot_longer(everything()) %>% 
  arrange(name) %>% 
  pander()
```

and models in this sample had between ...

```{r}
fuel %>% count(cyl)
```



Figure 2 suggests that both cylinders and weight were related to fuel economy

```{r}
fuel %>% 
  select(mpg, weight, cyl) %>% 
  ggplot(aes(weight, mpg, color=factor(cyl))) + 
  geom_smooth(method=lm, se=F, aes(group=1), color="black", size=.3) + 
  geom_point() + labs(color="Cylinders", x="Weight (kg)", y="Fuel Economy (mpg)") + 
  theme_classic()
  
```


To test whether cylinders and weight were independent predictors of fuel economy ...
we ran two linear regression models. Model A included only weight as a predictor; 
Model B included both weight and the number of cylinders.


```{r}
# fit lm models
modelA <- lm(grade ~ work_hours, data=studyhabits)
modelB <- lm(grade ~ work_hours + work_consistently, data=studyhabits)
```


Results are presented in Table X...

```{r}
# regression_table function is in the psydata package

regression_table(modelA) %>% pander(caption=modelA$call %>% deparse)
regression_table(modelB) %>% pander(caption=modelA$call %>% deparse)

```


Diagnostic plots for model B did not ...

```{r}
# this is a lazy shortcut to some diagnostics plots and produces more output 
# than we need; you could also use the techniques taught in the workshop using augment
modelB %>% plot()
```


Given these data—model B was ...

```{r}
# calculate the BF for model B vs model A

# first refit with lmBF
modA <- lmBF(grade ~ work_hours, data=studyhabits)
modB <- lmBF(grade ~ work_hours + work_consistently, data=studyhabits)

# calculate the BF for the comparison
modB/modA
```

