---
title: "Example data supplement"
output: html_document
authors: "Ben Whalley"
date: '2022-09-06'
---





### Overview

This document provides the analysis to support for Whalley et al, 2030: *"Disambiguating study duration and study consistency"*.

Each statement in that article requiring statistical analysis is copied below, and
followed by the necessary R code and output.

This supplement is NOT simply all the code used when analysing the data; it 
is an edited and selected subset of that. The analyses here are organised so that
they follow the logic of the paper, making it easier for readers to see where 
individual numbers in the text came from.



### Analyses

```{r}
# this tidies up the output somewhat
knitr::opts_chunk$set(message = F, warning=F)

library(tidyverse)
library(BayesFactor)
library(psydata)
library(pander)
```



```{r}
#model <- modelA
require(broom)

regression_table <- function(model, as_text=TRUE){
  x <- model %>% broom::tidy(conf.int=T) 
  
  z <- model %>% arm::standardize() %>% 
    broom::tidy()
  
  znum <- z %>% 
    mutate(term = str_replace_all(term, "z\\.", "")) %>% 
    transmute(term=term, β = estimate) %>% 
    left_join(x, ., by="term") %>% 
    select(Term=term, Estimate=estimate, `Lower`=conf.low, Upper=conf.high, β, p=p.value)
  
  r2 <- model %>% glance() %>% pull(adj.r.squared)
  znumr2 <- znum %>% 
    bind_rows(., tibble(Term="Adjusted R sq.", Estimate=r2)) 
  
  if(as_text){
    znumr2 %>% 
      mutate(`95% CI` = ifelse(!is.na(Lower), 
        glue::glue("[{sprintf('%.3f', Lower)}, {sprintf('%.3f', Upper)}]"), ""), 
        .after="Estimate") %>% 
      mutate_at(c("Estimate"), ~{ifelse(is.na(.x), "", sprintf("%.3f", .x) )}) %>% 
      mutate_if(is.numeric, ~{ifelse(is.na(.x), "", sprintf("%.2f", .x) )}) %>% 
      mutate_at(c("p"), ~ifelse(is.na(.x)|.x=="", "", apastats::round.p(.x))) %>% 
      select(-Lower, -Upper)
      
  } else {
    znumr2 %>% mutate_if(is.numeric, ~round(.x, 3))
  }
  
}
```



### Specific results from the article

Data were obtained for 300 students.

```{r}
# note we are using data from psydata package
# normally this would be the point at which you read in a csv or other data file
studyhabits %>% glimpse
```



Desxcriptive statistics for students:

```{r}
studyhabits %>% 
  pivot_longer(c(-unique_id, -female)) %>% 
  group_by(name) %>% 
  summarise_at("value", list(M=mean, Median = median, SD=sd, Min=min, Max=max)) %>% 
  rename(Name = name) %>% 
  mutate(Name = snakecase::to_title_case(Name))
  pander()
```


```{r}
studyhabits %>% 
  group_by(work_consistently) %>% count()
```


```{r}
studyhabits %>% 
  pivot_longer(c(work_hours, work_consistently)) %>% 
  ggplot(aes(value, grade)) + geom_jitter(alpha=.5) + geom_smooth(se=F) +
  facet_wrap(~snakecase::to_title_case(name), scales="free") + theme_minimal() + 
  labs(x="Score/Hours", y="Grade")

ggsave('studyhabs.pdf', width=4, height=2.5)
```



To test whether hours and consistency were independent predictors of grades, and to estimate the strength of these relationships… 
…we ran two linear regression models. Model A included only study hours as a predictor; Model B included both hours and the consistency. 

```{r}
# fit lm models
modelA <- lm(grade ~ work_hours, data=studyhabits)
modelB <- lm(grade ~ work_hours + work_consistently, data=studyhabits)
```

Results are presented in Table X...


```{r}
modelA %>% tidy %>% pander(digits=3)
modelA %>% glance %>% pander(digits=3)
modelA %>% confint %>% pander(digits=3)
modelA %>% arm::standardize(standardize.y=T) %>% pander(digits=3)
```


```{r}
modelB %>% tidy %>% pander(digits=3)
modelB %>% glance %>% pander(digits=3)
modelB %>% confint %>% pander(digits=3)
modelB %>% arm::standardize(standardize.y=T) %>% pander(digits=3)
```


Given these data...

```{r}
# calculate the BF for model B vs model A

# first refit with lmBF
modelAbf <- lmBF(grade ~ work_hours, data=studyhabits)
modelBbf <- lmBF(grade ~ work_hours + work_consistently, data=studyhabits)

# calculate the BF for the comparison
modelAbf
modelBbf/modelAbf
```


```{r}
library(rstanarm)



```

